---
title: "Arkansas Covid Daily Update - County"
author: "Rob Wells"
date: "8/29/2020"
output: pdf_document
---

# Daily Updating Arkansascovid.com County Data, Calculations

- **This notebook retrieves the daily API feed on the Austin Wilkins spreadsheet, performs all of the county calculations and rejoins it with the main DateMaster sheet.**

--**It does not handle the state and hospital calculations. That will come later**

### Retrieve Arkansas Covid Data from ADH FEED


- **The records come from Arkansascovid Google Drive**  
https://drive.google.com/drive/u/1/folders/1tsNmE6zeDdwpr6mVFacEXDOjn2RAjlJ3


- **Using this sheet: Date Master API testing**
https://docs.google.com/spreadsheets/d/16mdoWm0k3JUllm3InhoD_Jiq0lbArO_UGum8NMp9USU/edit#gid=0

--------------------------------------------------------------------
# Part 1: Fix the data structure

```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages("googledrive")
#install.packages("googlesheets4")
library(googlesheets4)
library(googledrive)
library(tidyr)
# vignette("basic-usage", package="googlesheets")
```



```{r}
url <- ("https://docs.google.com/spreadsheets/d/16mdoWm0k3JUllm3InhoD_Jiq0lbArO_UGum8NMp9USU/edit#gid=0")

covid <- read_sheet(url)
```

-**Google Pain in the Ass Workaround**
```{r}
covid <- rio::import("THE DOWNLOADED FILE HERE")
```

- **Clean Names**
```{r}

covid <- janitor::clean_names(covid)
```

**Examine the Data Fields**
```{r}
glimpse(covid)
```

```{r}
names(covid)

```

- **Extract Population**
```{r}
population <- rio::import("./Exercises/population.csv")
```


- **Each Day:** We join the new data, api, with the population and then perform the calculations

```{r}
daily <- covid %>% 
   inner_join(population, by="county")
head(daily)
```

# Part 2: MATH!   

- **Now We Do The Math**
- **Previous Date Calculations**

Create Temporary Table Two Days' Worth of Data
Sort Alphbetically and Run Calculations
You should get a df with 150 observations (two days' worth a data)
```{r}
temp <- daily %>% 
  filter(date > "2020-08-28") %>%  #INPUT A DATE FROM TWO DAYS AGO to get two days' of data
arrange(county) 
```

```{r}
head(temp) 
```

```{r}
glimpse(temp)
```

-**The Today-Yesterday Calculations**

    This calculates:
    Column K (New Cases Today)=J2-J3 
    Column N (Recovered Since Yesterday: =M2-M3)
    Column P (New Deaths Today: =O2-O3)
    Column R (New Tests Dashboard: =Q2-Q3)
    

```{r}
temp <- temp %>%
  mutate(new_cases_today = (positive-lead(positive))) %>%
  mutate(recovered_since_yesterday = (recovered-lead(recovered))) %>%
  mutate(new_deaths_today = (total_deaths-lead(total_deaths))) %>%
  mutate(new_tests_dashboard = (number_tested-lead(number_tested)))
```

- **IMPORTANT: FILTER TABLE TO TODAY'S RESULTS**
```{r}
temp <- temp %>% filter(date > "2020-08-27") #USE PRIOR DAY - SHOULD GET 75 ROWS IN Temp
```

**Fact Checked: OK**

Math on the 106-110
x <- temp %>% 
  select(date, county, positive, new_cases_today)
x
Math on the 106-110
x <- temp %>% 
  select(date, county, recovered, recovered_since_yesterday)
x

```{r}
glimpse(temp)
```

-**The Percentage Calculations**

    Column E (Cases/population = (positive / population) *100
    Column F (Tested/population = (Number tested / population) *100
    Column G (New Cases Today per 10k Population = (New Cases Today / population) *10000
    Column H (Active Cases per 10k = (current_infections / population) *10000
    Column T (% Positive Cumulative = (positive / Number tested) *100
    Column U (% Positive New to Dashboard = New Cases Today / New Tests Dashboard) * 100
    Column V (Closed = Recovered + Total Deaths)
    Column W (% Deaths vs. Recoveries = Total Deaths / Closed) * 100
    Column X (% Recoveries vs. Deaths = Recovered / Closed) * 100

```{r}
temp <- temp %>%
  mutate(cases_population = (positive / population)*100) %>% 
  mutate(tested_population = (number_tested / population)*100) %>% 
  mutate(new_cases_today_10k_pop = (new_cases_today/population)*10000) %>% 
  mutate(active_cases_10k_pop = (current_infections/population)*10000) %>% 
  mutate(pct_positive_cumulative = (positive/number_tested)*100) %>% 
  mutate(pct_positive_new_to_dashboard = (new_cases_today/new_tests_dashboard)*100) %>% 
  mutate(closed = (recovered + total_deaths)) %>% 
  mutate(pct_deaths_vs_recoveries = (total_deaths/closed)*100) %>% 
  mutate(pct_recoveries_vs_deaths = (recovered/closed)*100)
  
temp
```

```{r}
glimpse(temp)
```


**Fact Checked: OK**
Math on 142-150
# Column U (% Positive New to Dashboard = New Cases Today / New Tests Dashboard) * 100
x <- temp %>% 
  select(date, county, new_cases_today, new_tests_dashboard, pct_positive_new_to_dashboard)
x
- **IMPORTANT: Hospital and state calculations in a separate sheet**


align names in order of covid table
```{r}
temp <- temp %>% 
  select(county,date,population,cases_population,tested_population,new_cases_today_10k_pop,active_cases_10k_pop,current_infections,positive, new_cases_today, negative, recovered, recovered_since_yesterday, total_deaths, new_deaths_today, number_tested, new_tests_dashboard, pct_positive_cumulative, pct_positive_new_to_dashboard, closed, pct_deaths_vs_recoveries, pct_recoveries_vs_deaths)
names(temp)
```

-**This now matches Columns A-X in BeAll Sheet, DateMaster**

# Part #3: JOINING AND ARCHIVING   

-**Update the main sheet, archiving**
```{r}
#Create a standalone copy of the TODAY'S data with the calculations
Day_8_29_2020 <- temp
```


- **Retrieve BeAll from DateMaster**
```{r include=FALSE}
url2 <- ("https://docs.google.com/spreadsheets/d/1g-gkjJOr1sKAu6rZHG04XA5_fM_Ma0jLr5r24fwMqiA/edit#gid=1357388768")

beall <- read_sheet(url2, sheet="BeAll")
```

```{r}
beall <- janitor::clean_names(beall)
names(beall)

```


RENAME COLUMNS TO MATCH NEW FILE
```{r}
covid_temp <- beall %>% 
  #rename(NEW NAME = EXISTING NAME) %>% 
  rename(new_cases_today_10k_pop = new_cases_today_per_10k_population) %>% 
  rename(active_cases_10k_pop = active_cases_per_10k) %>% 
  rename(pct_positive_cumulative = percent_positive_cumulative) %>% 
  rename(pct_positive_new_to_dashboard = percent_positive_new_to_dashboard) %>% 
  rename(pct_deaths_vs_recoveries = percent_deaths_vs_recoveries) %>% 
  rename(pct_recoveries_vs_deaths = percent_recoveries_vs_deaths)
```    

- **This aligns DateMaster to the Day_8292020 file**

Data cleaning and normalization. Cut state (C), new case date corrected (S) column, hospital data (Y-AT)

- **Hospital data is a separate process**

```{r}

covid_temp <- covid_temp[ -c(3,19,25:46) ]
glimpse(covid_temp)
names(covid_temp)
```

-**Join Append the New and Existing**
```{r}
covid_update <- rbind(Day_8_29_2020,covid_temp)
write.csv(covid_update, "covid_update.csv")
write.csv(Day_8_29_2020, "Day_8_29_2020.csv")
```


**30**
