---
title: "Demographic Update Arkansas Covid"
author: "Rob Wells. Thanks to Mary Hennigan for assistance"
date: "10/13/2020"
output: pdf_document
---

# Demographic Daily Update Arkansascovid.com Calculations

# Part 1: Import Data, Clean It

```{r include=FALSE}
#install.packages("slider")
#install.packages("zoo")
#install.packages("gtools")
# vignette("basic-usage", package="googlesheets")
#install.packages("googledrive")
#install.packages("googlesheets4")
library(tidyverse)
library(janitor)
library(lubridate)
#library(googlesheets4)
#library(googledrive)
library(tidyr)
library(jsonlite)
library(gtools)
library(zoo)  
library(reshape2)
library(slider)

```

## Retrieve State Arkansas Covid Data from ADH FEED
```{r}
#New County json feed
#38 Variables
z <- fromJSON('https://services.arcgis.com/PwY9ZuZRDiI5nXUB/ArcGIS/rest/services/UPDATED_ADH_COVID19_STATE_METRICS/FeatureServer/0/query?where=0%3D0&objectIds=&time=&resultType=none&outFields=*&returnIdsOnly=false&returnUniqueIdsOnly=false&returnCountOnly=false&returnDistinctValues=false&cacheHint=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&having=&resultOffset=&resultRecordCount=&sqlFormat=none&f=pjson&token=')

df <- z[["features"]][["attributes"]]
#print's today's date as a date field
df$date <- Sys.Date()
#file.exists(filename)
#austin help on that
#if file exists = check if current data is different and if that is true then append today's data set to master dataset and resave
today_demo <- df
```

#This is the API feed. It is called "today_demo" for demographics. Get it?
```{r}
today_demo <- janitor::clean_names(today_demo)
#if you miss a date, uncomment below and use this
#today_demo$date <- ("2020-10-23")
```


#Set Dates
```{r}
#today's date
today <- Sys.Date()

#NOTE: IF YOU ARE RUNNING THIS A DAY LATE, USE THIS CODE TO WORK PROPERLY
#today <- Sys.Date()-1
#today_demo$mydate <-"2020-09-22- THE OLD DATE...."

#yesterday's date
yesterday <- (today-1)

```

#Cleaning Names, Converting to Numeric Variables
```{r}
#today_demo <- today_demo [ -c(1) ]
today_demo [2:58] <- lapply(today_demo [2:58], as.numeric)
today_demo$date <- as.Date(today_demo$date)
glimpse(today_demo)
```

#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#AGE calculations
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#Today's data with age
```{r}
today_age <- today_demo %>% 
  select(date, age_1, age_2, age_3, age_4, age_5, age_6, age_7, age_8, age_9, d_age_1,d_age_2, d_age_3, d_age_4, d_age_5,  d_age_6,  d_age_7,  d_age_8,  d_age_9)
```

#Rename

```{r}
today_age2 <- today_age

colnames(today_age2)[2:19] <-c("cases_0_10", "cases_11_17", "cases_18_24", "cases_25_34", "cases_35_44", "cases_45_54", "cases_55_64", "cases_65", "cases_unknown", "deaths_0_10", "deaths_11_17", "deaths_18_24", "deaths_25_34", "deaths_35_44", "deaths_45_54", "deaths_55_64", "deaths_65", "deaths_unknown")
today_age2$date <- as.Date(today_age2$date)
glimpse(today_age2)
```

#Today_yesterday
#Import age master file
```{r}
age3 <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/age_master.csv")

age3 <- janitor::clean_names(age3)
age3$date <- as.Date(age3$date)
glimpse(age3)
```

```{r}
yesterday_age <- age3 %>% 
   select("date", "cases_0_10", "cases_11_17", "cases_18_24", "cases_25_34", "cases_35_44", "cases_45_54", "cases_55_64", "cases_65", "cases_unknown", "deaths_0_10", "deaths_11_17", "deaths_18_24", "deaths_25_34", "deaths_35_44", "deaths_45_54", "deaths_55_64", "deaths_65", "deaths_unknown") %>% 
  filter(date==yesterday)

#if you miss a day, fill this out and run instead
#filter(date == "2020-10-19") %>% 

 yesterday_age$date <- as.Date(yesterday_age$date)
 glimpse(yesterday_age)
```



#This creates a df with 2 observations (two days' worth a data)
```{r}
twodays <- rbind(today_age2, yesterday_age)
glimpse(twodays)

```


-**The Today-Yesterday Calculations**

```{r}
twodays <- twodays %>%
  mutate(new_cases_0_10 = (cases_0_10-lead(cases_0_10))) %>% 
  mutate(new_cases_11_17 = (cases_11_17-lead(cases_11_17))) %>% 
  mutate(new_cases_18_24 = (cases_18_24-lead(cases_18_24))) %>% 
  mutate(new_cases_25_34 = (cases_25_34-lead(cases_25_34))) %>%  
  mutate(new_cases_35_44 = (cases_35_44-lead(cases_35_44))) %>%  
  mutate(new_cases_45_54 = (cases_45_54-lead(cases_45_54))) %>%
  mutate(new_cases_55_64 = (cases_55_64-lead(cases_55_64))) %>%
  mutate(new_cases_65 = (cases_65-lead(cases_65))) %>% 
  mutate(new_deaths_0_10 = (deaths_0_10-lead(deaths_0_10))) %>% 
  mutate(new_deaths_11_17 = (deaths_11_17-lead(deaths_11_17))) %>% 
  mutate(new_deaths_18_24 = (deaths_18_24-lead(deaths_18_24))) %>% 
  mutate(new_deaths_25_34 = (deaths_25_34-lead(deaths_25_34))) %>%  
  mutate(new_deaths_35_44 = (deaths_35_44-lead(deaths_35_44))) %>%  
  mutate(new_deaths_45_54 = (deaths_45_54-lead(deaths_45_54))) %>%
  mutate(new_deaths_55_64 = (deaths_55_64-lead(deaths_55_64))) %>%
  mutate(new_deaths_65 = (deaths_65-lead(deaths_65))) 
```

#Creates one day with Today's calculations only
```{r}
today_age_calcs <- twodays %>% 
    filter(date ==today)
  #if you miss a day, fill this out and run instead
  #filter(date == "2020-10-14") 

today_age_calcs <- janitor::clean_names(today_age_calcs)
glimpse(today_age_calcs)
```

```{r}
age4 <- smartbind(age3, today_age_calcs)

age4 <- age4 %>% 
  arrange(desc(date))

#remove first two columns of garbage index if needed
#age4 <- age4[ -c(1,2) ]

write.csv(age4, "age_master.csv")
```

#New Cases Today Table for Web
```{r}
age_new_cases <- age4 %>% 
  select(date, new_cases_0_10, new_cases_11_17, new_cases_18_24, new_cases_25_34, new_cases_35_44, new_cases_45_54, new_cases_55_64, new_cases_65) %>% 
  filter(date == today)

#flip
age_new_cases <- gather(age_new_cases, key = "age", value = "new_cases", new_cases_0_10, new_cases_11_17, new_cases_18_24, new_cases_25_34, new_cases_35_44, new_cases_45_54, new_cases_55_64, new_cases_65)

age_new_cases
write.csv(age_new_cases, "age_new_cases.csv")

```

#New Deaths Today Table for Web
```{r}
age_new_deaths <- age4 %>% 
  select(date, new_deaths_0_10, new_deaths_11_17, new_deaths_18_24,  new_deaths_25_34, new_deaths_35_44, new_deaths_45_54, new_deaths_55_64, new_deaths_65) %>% 
  filter(date == today)

#flip
age_new_deaths <- gather(age_new_deaths, key = "age", value = "new_deaths", new_deaths_0_10, new_deaths_11_17, new_deaths_18_24, new_deaths_25_34, new_deaths_35_44, new_deaths_45_54, new_deaths_55_64, new_deaths_65)

age_new_deaths
write.csv(age_new_deaths, "age_new_deaths.csv")

```



#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#RACE calculations
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#Today's data with race
```{r}
today_race <- today_demo %>% 
  select(date, black, white, na, asian, pi, hispanic, multi_race, other_race, unk_race, d_black, d_white, d_na, d_asian, d_pi, d_hispanic, d_multi_race, d_other_race, d_unk_race)
```

#Rename

```{r}
today_race2 <- today_race
colnames(today_race2)[2:19] <-c("cases_african_american", "cases_white", "cases_native_american", "cases_asian", "cases_pacific_islander", "cases_hispanic", "cases_multi_racial", "cases_other_race", "cases_unknown_race", "deaths_african_american", "deaths_white", "deaths_native_american", "deaths_asian", "deaths_pacific_islander", "deaths_hispanic", "deaths_multi_racial", "deaths_other_race", "deaths_unknown_race")


today_race2$date <- as.Date(today_race2$date)

today_race2 <- janitor::clean_names(today_race2)

glimpse(today_race2)
```

#Today_yesterday
#Import the master file for race from GitHub

```{r}
race3 <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/race_main_file.csv")
race3 <- janitor::clean_names(race3)
glimpse(race3)
```


```{r}
race3 <- as.data.frame(race3)
yesterday_race <- race3 %>% 
  select("date", "cases_african_american", "cases_white", "cases_native_american", "cases_asian", "cases_pacific_islander", "cases_hispanic", "cases_multi_racial", "deaths_african_american", "deaths_white", "deaths_native_american", "deaths_asian", "deaths_pacific_islander", "deaths_hispanic", "deaths_multi_racial", "total_state_population_african_american", "total_state_population_asian", "total_state_population_hispanic", "total_state_population_multi_racial", "total_state_population_native_american",
"total_state_population_pacific_islander", "total_state_population_white",  "total_state_population_other", "total_state_population_race_or_ethnicity_unknown", "pct_state_population_african_american",  "pct_state_population_asian", "pct_state_population_hispanic", "pct_state_population_multi_racial",  "pct_state_population_native_american", "pct_state_population_pacific_islander", 
"pct_state_population_white", "pct_state_population_other", "pct_state_population_race_or_ethnicity_unknown") %>% 
    filter(date == yesterday)


#if you miss a day, fill this out and run instead
# filter(date == "2020-10-13") %>% 

yesterday_race$date <- as.Date(yesterday_race$date)
glimpse(yesterday_race)
```

#Create a population table
```{r}
pop <- race3 %>% 
  select("date", "total_state_population_african_american", "total_state_population_asian", "total_state_population_hispanic", "total_state_population_multi_racial", "total_state_population_native_american",
"total_state_population_pacific_islander", "total_state_population_white",  "total_state_population_other", "total_state_population_race_or_ethnicity_unknown", "pct_state_population_african_american",  "pct_state_population_asian", "pct_state_population_hispanic", "pct_state_population_multi_racial",  "pct_state_population_native_american", "pct_state_population_pacific_islander", 
"pct_state_population_white", "pct_state_population_other", "pct_state_population_race_or_ethnicity_unknown") %>% 
  filter(date == yesterday) 

#if you miss a day, fill this out and run instead
#filter(date == "2020-10-13") %>% 
#
#Change to today's date
pop$date <- today
#if you miss a day, fill this out and run instead
#pop$date <- "2020-10-14"

pop$date <- as.Date(pop$date)
glimpse(pop)
```


```{r}
today_race2 <- today_race2 %>% 
   left_join(pop, by="date")
head(today_race2)
```

#You should get a df with 2 observations (two days' worth a data)
```{r}
twodays2 <- smartbind(today_race2, yesterday_race)
twodays2$date <- as.Date(twodays2$date)
glimpse(twodays2)

```

-**The Today-Yesterday Calculations**

```{r}
twodays3 <- twodays2 %>%
  mutate(New_cases_african_american = (cases_african_american-lead(cases_african_american))) %>% 
  mutate(New_cases_white = (cases_white-lead(cases_white))) %>% 
   mutate(New_cases_native_american = (cases_native_american-lead(cases_native_american))) %>%
   mutate(New_cases_asian = (cases_asian-lead(cases_asian))) %>%
   mutate(New_cases_pacific_islander = (cases_pacific_islander-lead(cases_pacific_islander))) %>%
   mutate(New_cases_hispanic = (cases_hispanic-lead(cases_hispanic))) %>%
   mutate(New_cases_multi_racial = (cases_multi_racial-lead(cases_multi_racial))) %>%
   mutate(New_deaths_african_american = (deaths_african_american-lead(deaths_african_american))) %>%
   mutate(New_deaths_white = (deaths_white-lead(deaths_white))) %>%
   mutate(New_deaths_native_american = (deaths_native_american-lead(deaths_native_american))) %>%
   mutate(New_deaths_asian = (deaths_asian-lead(deaths_asian))) %>%
   mutate(New_deaths_pacific_islander = (deaths_pacific_islander-lead(deaths_pacific_islander))) %>%
   mutate(New_deaths_hispanic = (deaths_hispanic-lead(deaths_hispanic))) %>%
   mutate(New_deaths_multi_racial = (deaths_multi_racial-lead(deaths_multi_racial))) 
```

#Percentage Calculations 
```{r}
twodays3 <- twodays3 %>%
mutate(cases_per_10_000_african_american = (cases_african_american / total_state_population_african_american)*10000) %>% 
mutate(cases_per_10_000_asian = (cases_asian / total_state_population_asian)*10000) %>% 
mutate(cases_per_10_000_multi_racial = (cases_multi_racial / total_state_population_multi_racial)*10000) %>%
mutate(cases_per_10_000_pacific_islander = (cases_pacific_islander / total_state_population_pacific_islander)*10000) %>%
mutate(cases_per_10_000_hispanic = (cases_hispanic / total_state_population_hispanic)*10000) %>%
mutate(cases_per_10_000_native_american = (cases_native_american / total_state_population_native_american)*10000) %>%
mutate(cases_per_10_000_white = (cases_white / total_state_population_white)*10000) %>%
mutate(deaths_per_10_000_african_american = (deaths_african_american / total_state_population_african_american)*10000) %>% 
mutate(deaths_per_10_000_asian = (deaths_asian / total_state_population_asian)*10000) %>% 
mutate(deaths_per_10_000_multi_racial = (deaths_multi_racial / total_state_population_multi_racial)*10000) %>%
mutate(deaths_per_10_000_pacific_islander = (deaths_pacific_islander / total_state_population_pacific_islander)*10000) %>%
mutate(deaths_per_10_000_hispanic = (deaths_hispanic / total_state_population_hispanic)*10000) %>%
mutate(deaths_per_10_000_native_american = (deaths_native_american / total_state_population_native_american)*10000) %>%
mutate(deaths_per_10_000_white = (deaths_white / total_state_population_white)*10000)

```


#Merges the Today's calculations with the master file
```{r}
today_race_calcs <- twodays3 %>% 
  filter(date ==today)
#if you miss a day, fill this out and run instead
#filter(date == "2020-10-14")
  
  
today_race_calcs <- janitor::clean_names(today_race_calcs)
glimpse(today_race_calcs)
```

#Writing the Final Updated File


```{r}
race4 <- smartbind(race3, today_race_calcs)

race4 <- race4 %>% 
  arrange(desc(date))

#remove first two columns of garbage index if needed
#race4 <- race4[ -c(1,2) ]

write.csv(race4, "race_main_file.csv")
```
#New Cases Today Table for Web
```{r}
x <- race4 %>% 
  select(date, new_cases_african_american, new_cases_asian, new_cases_hispanic, new_cases_multi_racial, new_cases_native_american, new_cases_other, new_cases_race_or_ethnicity_unknown, new_cases_white) %>% 
  filter(date == today)

#flip
race_new_cases <- gather(x, key = "race", value = "new_cases", new_cases_african_american, new_cases_asian, new_cases_hispanic, new_cases_multi_racial, new_cases_native_american, new_cases_other, new_cases_race_or_ethnicity_unknown, new_cases_white)

race_new_cases
write.csv(race_new_cases, "race_new_cases.csv")
```

#New Deaths Today Table for Web
```{r}
race_new_deaths <- race4 %>% 
  select(date, new_deaths_african_american, new_deaths_asian, new_deaths_hispanic, new_deaths_multi_racial, new_deaths_native_american, new_deaths_other, new_deaths_race_or_ethnicity_unknown, new_deaths_white) %>% 
  filter(date == today)

#flip
race_new_deaths <- gather(race_new_deaths, key = "race", value = "new_cases", new_deaths_african_american, new_deaths_asian, new_deaths_hispanic, new_deaths_multi_racial, new_deaths_native_american, new_deaths_other, new_deaths_race_or_ethnicity_unknown, new_deaths_white)

race_new_deaths
write.csv(race_new_deaths, "race_new_deaths.csv")
```





#----------------------------------------------------------------------------------------------------#
#Gender
#----------------------------------------------------------------------------------------------------#
#Today's data 
```{r}
today_gender <- today_demo %>% 
  select(date, male, female, unk_gender)
```


#Rename

```{r}
today_gender2 <- today_gender

colnames(today_gender2)[2:4] <-c("cases_male", "cases_female", "cases_gender_unknown")

today_gender2$date <- as.Date(today_gender2$date)

today_gender2 <- janitor::clean_names(today_gender2)

glimpse(today_gender2)
```

#Today_yesterday
#Import the master file for gender

```{r}
#gender_master <- rio::import("gender_master_10_11.csv")
gender_master <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/gender_master.csv")

gender_master <- janitor::clean_names(gender_master)

yesterday_gender <- gender_master %>% 
   filter(date == yesterday)

yesterday_gender <- gender_master %>% 
  select(date, cases_male, cases_female, cases_gender_unknown)

glimpse(yesterday_gender)
```

#You should get a df with 2 observations (two days' worth a data)
```{r}
twodays3 <- smartbind(today_gender2, yesterday_gender)
glimpse(twodays3)

```

-**The Today-Yesterday Calculations**

```{r}
gender2 <- twodays3 %>%
  mutate(new_cases_male = (cases_male -lead(cases_male))) %>% 
  mutate(new_cases_female = (cases_female-lead(cases_female))) %>% 
  mutate(new_cases_gender_unknown = (cases_gender_unknown-lead(cases_gender_unknown)))

glimpse(gender2)

write.csv(gender2, "gender_master.csv")
```

#New Cases Today Table for Web
```{r}
gender_new_cases <- gender2 %>% 
  select(date, new_cases_male, new_cases_female, new_cases_gender_unknown) %>% 
  filter(date == today)

#flip
gender_new_cases <- gather(gender_new_cases, key = "gender", value = "new_cases", new_cases_male, new_cases_female, new_cases_gender_unknown)

gender_new_cases
write.csv(gender_new_cases, "gender_new_cases.csv")
```



#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
# NOTES BELOW
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#

#Build this 
#----------------------------------------------------------------------------------------------------#
#Facility
#----------------------------------------------------------------------------------------------------#

```{r}
(facility2, "facility_10_13.csv")
```

#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#

```{r}
glimpse(today_demo)
```

#Build this 



#MISC

```{r}
glimpse(today_demo)
today_demo_10_12 <- today_demo
write.csv(today_demo, "today_demo_10_12.csv")
```




#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
# STOP!!!!!
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#


#This section should be the only time we run this to break down the old spreadsheet format into machine-readable language
-**Import Demographic Data for Cleaning**
```{r}
demo_table <- rio::import("/Users/rswells/Dropbox/Classes/Data Fall 2020/ArkansasCovid/Marys_Demo_Table_10_11.csv")

demo_table$date <- lubridate::mdy(demo_table$date)
demo_table <- janitor::clean_names(demo_table)
glimpse(demo_table)
```

#age 
```{r}
age <- demo_table %>% 
  filter(detail == "Age")

age2 <- age %>% 
  select(date, sub_detail, cases, new_cases, deaths, new_deaths)

age3 <- age2 %>% 
  pivot_wider(names_from = sub_detail, values_from = c(cases, new_cases, deaths, new_deaths))
write.csv(age3, "age_10_13.csv")
```
#One Time Building of the Age table
```{r}
# colnames(age3)[29] <- "deaths_11_17"
# age3B <- age3 [ -c(1:2) ]
# names(age3B)
# 
# age3B <- age3B %>% 
#    select("date", "cases_0_10", "cases_11_17", "cases_18_24", "cases_25_34", "cases_35_44", "cases_45_54", "cases_55_64", "cases_65", "cases_unknown", "deaths_0_10", "deaths_11_17", "deaths_18_24", "deaths_25_34", "deaths_35_44", "deaths_45_54", "deaths_55_64", "deaths_65", "deaths_unknown", "cases_0_17", "cases_25_44", "cases_45_64", "deaths_0_17", "deaths_25_44", "deaths_45_64")
# names(age3B)
# 
# age3D <- age3B %>% 
#   mutate(new_cases_0_10 = (cases_0_10-lead(cases_0_10))) %>% 
#   mutate(new_cases_11_17 = (cases_11_17-lead(cases_11_17))) %>% 
#   mutate(new_cases_18_24 = (cases_18_24-lead(cases_18_24))) %>% 
#   mutate(new_cases_25_34 = (cases_25_34-lead(cases_25_34))) %>%  
#   mutate(new_cases_35_44 = (cases_35_44-lead(cases_35_44))) %>%  
#   mutate(new_cases_45_54 = (cases_45_54-lead(cases_45_54))) %>%
#   mutate(new_cases_55_64 = (cases_55_64-lead(cases_55_64))) %>%
#   mutate(new_cases_65 = (cases_65-lead(cases_65))) %>% 
#   mutate(new_deaths_0_10 = (deaths_0_10-lead(deaths_0_10))) %>% 
#   mutate(new_deaths_11_17 = (deaths_11_17-lead(deaths_11_17))) %>% 
#   mutate(new_deaths_18_24 = (deaths_18_24-lead(deaths_18_24))) %>% 
#   mutate(new_deaths_25_34 = (deaths_25_34-lead(deaths_25_34))) %>%  
#   mutate(new_deaths_35_44 = (deaths_35_44-lead(deaths_35_44))) %>%  
#   mutate(new_deaths_45_54 = (deaths_45_54-lead(deaths_45_54))) %>%
#   mutate(new_deaths_55_64 = (deaths_55_64-lead(deaths_55_64))) %>%
#   mutate(new_deaths_65 = (deaths_65-lead(deaths_65))) %>% 
#   mutate(new_cases_0_17 = (cases_0_17-lead(cases_0_17))) %>% 
#   mutate(new_cases_25_44 = (cases_25_44-lead(cases_25_44))) %>%
#   mutate(new_cases_45_64 = (cases_45_64-lead(cases_45_64))) %>%
#   mutate(new_deaths_0_17 = (deaths_0_17-lead(deaths_0_17))) %>%
#   mutate(new_deaths_25_44 = (deaths_25_44-lead(deaths_25_44))) %>%
#   mutate(new_deaths_45_64 = (deaths_45_64-lead(deaths_45_64)))
# write.csv(age3D, "age_update_10_12.csv")
```

#facility
```{r}
facility <- demo_table %>% 
  filter(detail == "Facility Type")

#clean
facility <- facility %>% 
  select(date, sub_detail, cases, new_cases, deaths, new_deaths)
glimpse(facility)

facility2 <- facility %>% 
  pivot_wider(names_from = sub_detail, values_from = c(cases, new_cases, deaths, new_deaths))

write_csv(facility2, "facility_10_13.csv")
```

#race
```{r}
race <- demo_table %>% 
  filter(detail == "Race or Ethnicity")
glimpse(race)
#clean

race2 <- race %>% 
  select(date, sub_detail, cases, new_cases, deaths, new_deaths, total_state_population, pct_state_population, cases_per_10_000, deaths_per_10_000)
glimpse(race2)

race3 <- race2 %>% 
  pivot_wider(names_from = sub_detail, values_from = c(cases, new_cases, deaths, new_deaths, total_state_population, pct_state_population, cases_per_10_000, deaths_per_10_000))

glimpse(race3)

write.csv(race3, "race_10_13.csv")

```

```{r}
glimpse(gender_master)
```

#gender create the table
```{r}
gender <- demo_table %>% 
  filter(detail == "Gender")

gender2 <- gender %>% 
  select(date, sub_detail, cases, new_cases)
glimpse(gender2)

write.csv(gender2, "gender_10_13.csv")
#clean

gender_master2 <- gender_master %>% 
  select(date, sub_detail, cases, new_cases)

gender_master2 <- gender_master2 %>% 
  pivot_wider(names_from = sub_detail, values_from = c(cases, new_cases))

write.csv(gender_master2, "gender_master_10_11.csv")
```

#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
# END OF ONE TIME SECTION
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#




-**30**


