---
title: "Demographic Update Arkansas Covid"
author: "Rob Wells and Mary Hennigan, Translation by Rachell Sanchez-Smith"
date: "11/30/2020"
output: pdf_document
---

#This is New
#This is New
#Create the following folders in your working directory in R
Demographics
MasterData
SpanishData
HomePageData



# Demographic Daily Update Arkansascovid.com Calculations

# Part 1: Import Data, Clean It

```{r include=FALSE}
#install.packages("slider")
#install.packages("zoo")
#install.packages("gtools")
# vignette("basic-usage", package="googlesheets")
#install.packages("googledrive")
#install.packages("googlesheets4")
library(tidyverse)
library(janitor)
library(lubridate)
#library(googlesheets4)
#library(googledrive)
library(tidyr)
library(jsonlite)
library(gtools)
library(zoo)  
library(reshape2)
library(slider)

```

## Retrieve State Arkansas Covid Data from ADH FEED
```{r}
#New County json feed
#38 Variables
z <- fromJSON('https://services.arcgis.com/PwY9ZuZRDiI5nXUB/ArcGIS/rest/services/UPDATED_ADH_COVID19_STATE_METRICS/FeatureServer/0/query?where=0%3D0&objectIds=&time=&resultType=none&outFields=*&returnIdsOnly=false&returnUniqueIdsOnly=false&returnCountOnly=false&returnDistinctValues=false&cacheHint=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&having=&resultOffset=&resultRecordCount=&sqlFormat=none&f=pjson&token=')

df <- z[["features"]][["attributes"]]
#print's today's date as a date field
df$date <- Sys.Date()
#file.exists(filename)
#austin help on that
#if file exists = check if current data is different and if that is true then append today's data set to master dataset and resave
today_demo <- df
```

#This is the API feed. It is called "today_demo" for demographics. Get it?
```{r}
today_demo <- janitor::clean_names(today_demo)
#if you miss a date, uncomment below and use this
#today_demo$date <- ("2020-11-01")
```


#Set Dates
```{r}
#today's date
today <- Sys.Date()

#NOTE: IF YOU ARE RUNNING THIS A DAY LATE, USE THIS CODE TO WORK PROPERLY
#today <- Sys.Date()-1
#today_demo$mydate <-"2020-09-22- THE OLD DATE...."

#yesterday's date
yesterday <- (today-1)


glimpse(today_demo)
```

#Cleaning Names, Converting to Numeric Variables
```{r}
#today_demo <- today_demo [ -c(1) ]
today_demo [2:58] <- lapply(today_demo [2:58], as.numeric)
today_demo$date <- as.Date(today_demo$date)
glimpse(today_demo)
```

#Rename Variables
```{r}
colnames(today_demo)[1:59] <-c("Objectid","Fips","Black", "White", "Na", 
 "Asian", "Pi","Other_Race","Unk_Race","Multi_Race", 
 "Nonhispanic", "Hispanic","Deaths_Black", "Deaths_White", "Deaths_Na", 
"Deaths_Asian", "Deaths_Pi","Deaths_Other_Race","Deaths_Unk_Race","Deaths_Multi_Race", 
"Cases_0_10", "Cases_11_17", "Cases_18_24", "Cases_25_34", "Cases_35_44", "Deaths_0_10"   
, "Deaths_11_17", "Deaths_18_24", "Deaths_25_34", "Deaths_35_44", "Male","Female","Unk_Gender","Diabetes","Cardiac_Disease",
"Hypertension","Cpd", "Ckd", "Immunocompromised", "Healthcare", 
"Nursing_Home","Jailed","Positives", "Deaths_Nursing_Home","Deaths_Jailed", 
"Deaths", "Deaths_Hispanic", "Deaths_Nonhispanic", "Cases_45_54", "Deaths_45_54", "Cases_55_64", "Deaths_55_64", "Cases_65", "Deaths_65", 
"Noncovid_Death", "Pos_24Hr", "Cases_Unknown", "Deaths_Unknown","Date")

glimpse(today_demo)
```

#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#AGE calculations
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#Today's data with age
```{r}
today_age <- today_demo %>% 
  select("Date", 
 "Cases_0_10"   
, "Cases_11_17" 
, "Cases_18_24" 
, "Cases_25_34" 
, "Cases_35_44" 
, "Deaths_0_10"    
, "Deaths_11_17"    
, "Deaths_18_24"    
, "Deaths_25_34"    
, "Deaths_35_44"
, "Cases_45_54" 
, "Deaths_45_54"    
, "Cases_55_64" 
, "Deaths_55_64"              
, "Cases_65"                               
, "Deaths_65"                        
, "Cases_Unknown"                               
, "Deaths_Unknown" )
glimpse(today_age)
```

#Today_yesterday
#Import age master file
```{r}
age3 <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/age_master.csv")
#age3 <- rio::import("age_master.csv")

#age3 <- janitor::clean_names(age3)
#age3$date <- mdy(age3$date)
#age3 <- age3[ -c(1,2) ]
glimpse(age3)
#write.csv(age3, "age3.csv")
```

```{r}
yesterday_age <- age3 %>% 
   select("Date", "Cases_0_10", "Cases_11_17", "Cases_18_24", "Cases_25_34", "Cases_35_44", "Cases_45_54", "Cases_55_64", "Cases_65", "Cases_Unknown", "Deaths_0_10", "Deaths_11_17", "Deaths_18_24", "Deaths_25_34", "Deaths_35_44", "Deaths_45_54", "Deaths_55_64", "Deaths_65", "Deaths_Unknown") %>% 
  filter(Date==yesterday)

#if you miss a day, fill this out and run instead
#filter(date == "2020-10-19") %>% 

 yesterday_age$Date <- as.Date(yesterday_age$Date)
 glimpse(yesterday_age)
```



#This creates a df with 2 observations (two days' worth a data)
```{r}
twodays <- rbind(today_age, yesterday_age)
glimpse(twodays)

```


-**The Today-Yesterday Calculations**

```{r}
twodays <- twodays %>%
   mutate(New_Cases_0_10 = (Cases_0_10-lead(Cases_0_10))) %>% 
  mutate(New_Cases_11_17 = (Cases_11_17-lead(Cases_11_17))) %>% 
  mutate(New_Cases_18_24 = (Cases_18_24-lead(Cases_18_24))) %>% 
  mutate(New_Cases_25_34 = (Cases_25_34-lead(Cases_25_34))) %>%  
  mutate(New_Cases_35_44 = (Cases_35_44-lead(Cases_35_44))) %>%  
  mutate(New_Cases_45_54 = (Cases_45_54-lead(Cases_45_54))) %>%
  mutate(New_Cases_55_64 = (Cases_55_64-lead(Cases_55_64))) %>%
  mutate(New_Cases_65 = (Cases_65-lead(Cases_65))) %>% 
  mutate(New_Deaths_0_10 = (Deaths_0_10-lead(Deaths_0_10))) %>% 
  mutate(New_Deaths_11_17 = (Deaths_11_17-lead(Deaths_11_17))) %>% 
  mutate(New_Deaths_18_24 = (Deaths_18_24-lead(Deaths_18_24))) %>% 
  mutate(New_Deaths_25_34 = (Deaths_25_34-lead(Deaths_25_34))) %>%  
  mutate(New_Deaths_35_44 = (Deaths_35_44-lead(Deaths_35_44))) %>%  
  mutate(New_Deaths_45_54 = (Deaths_45_54-lead(Deaths_45_54))) %>%
  mutate(New_Deaths_55_64 = (Deaths_55_64-lead(Deaths_55_64))) %>%
  mutate(New_Deaths_65 = (Deaths_65-lead(Deaths_65))) 
```

#Creates one day with Today's calculations only
```{r}
today_age_calcs <- twodays %>% 
    filter(Date ==today)
  #if you miss a day, fill this out and run instead
  #filter(date == "2020-10-14") 

#today_age_calcs <- janitor::clean_names(today_age_calcs)
glimpse(today_age_calcs)
```

```{r}
age_master <- smartbind(age3, today_age_calcs)

age_master <- age_master %>% 
  arrange(desc(Date))

#age_master <- slice(age_master, -c(1))
#remove first two columns of garbage index if needed
#age_master <- age_master[ -c(1) ]

write.csv(age_master, file = "Demographics/age_master.csv")
glimpse(age_master)
```

#Create age_FLOURISH_master
#Rename variables 
```{r}
age_master <- age_master %>% 
  filter(Date >= "2020-09-13")

#age_master <- age_master[ -c(1) ]
write.csv(age_master, file = "MasterData/age_FLOURISH_master.csv")
write.csv(age_master, file = "Demographics/age_FLOURISH_master.csv")

```


#New Cases Today Table for Flourish
```{r}
age_new_cases <- age_master %>% 
  select(Date, New_Cases_0_10, New_Cases_11_17, New_Cases_18_24, New_Cases_25_34, New_Cases_35_44, New_Cases_45_54, New_Cases_55_64, New_Cases_65) %>% 
  filter(Date == today) %>% 
  distinct()

#flip
age_new_cases <- gather(age_new_cases, key = "Age", value = "New_Cases", New_Cases_0_10, New_Cases_11_17, New_Cases_18_24, New_Cases_25_34, New_Cases_35_44, New_Cases_45_54, New_Cases_55_64, New_Cases_65)

age_new_cases
write.csv(age_new_cases, file = "Demographics/age_new_cases.csv")

#Spanish translation

age_new_cases_spanish <- age_new_cases %>% 
  rename(Nuevos_Casos = New_Cases, Grupo_de_Edad = Age, Fecha = Date) 

age_new_cases_spanish$Grupo_de_Edad <- str_replace_all(age_new_cases_spanish$Grupo_de_Edad, pattern=fixed('New_Cases_'), replacement=fixed('') )


write.csv(age_new_cases_spanish, file = "SpanishData/age_new_cases_spanish.csv")
age_new_cases_spanish
```

#New Deaths Today Table for Flourish
```{r}
age_new_deaths <- age_master %>% 
  select(Date, New_Deaths_0_10, New_Deaths_11_17, New_Deaths_18_24,  New_Deaths_25_34, New_Deaths_35_44, New_Deaths_45_54, New_Deaths_55_64, New_Deaths_65) %>% 
  filter(Date == today) %>% 
  distinct()

#flip
age_new_deaths <- gather(age_new_deaths, key = "Age", value = "New_Deaths", New_Deaths_0_10, New_Deaths_11_17, New_Deaths_18_24,  New_Deaths_25_34, New_Deaths_35_44, New_Deaths_45_54, New_Deaths_55_64, New_Deaths_65)

age_new_deaths
write.csv(age_new_deaths, file = "Demographics/age_new_deaths.csv")

#Spanish translation

age_new_deaths_spanish <- age_new_deaths %>% 
  rename(Nuevas_Muertes = New_Deaths, Grupo_de_Edad = Age, Fecha = Date) 

age_new_deaths_spanish$Grupo_de_Edad <- str_replace_all(age_new_deaths_spanish$Grupo_de_Edad, pattern=fixed('New_Deaths_'), replacement=fixed('') )

age_new_deaths_spanish

write.csv(age_new_deaths_spanish, file = "SpanishData/age_new_deaths_spanish.csv")
```
#Combine age_new_deaths and age_new_cases

```{r}
#strip the race name down
age_new_cases$NEW <- str_replace_all(age_new_cases$Age, pattern=fixed('New_Cases_'), replacement=fixed('Age_'))
age_new_deaths$NEW <- str_replace_all(age_new_deaths$Age, pattern=fixed('New_Deaths_'), replacement=fixed('Age_'))

```




```{r}
age_new_cases_deaths <- age_new_cases %>% 
   inner_join(age_new_deaths, by=c("NEW", "Date"))

age_new_cases_deaths <- age_new_cases_deaths %>% 
  select(Date, NEW, New_Cases, New_Deaths)

colnames(age_new_cases_deaths)[2] <- "Age"
head(age_new_cases_deaths)

write.csv(age_new_cases_deaths, file = "Demographics/age_new_cases_deaths.csv")

#Spanish translation

age_new_cases_deaths_spanish <- age_new_cases_deaths %>% 
  rename(Nuevas_Muertes = New_Deaths, Grupo_de_Edad = Age, Fecha = Date, Nuevos_Casos = New_Cases) 

age_new_cases_deaths_spanish$Grupo_de_Edad <- str_replace_all(age_new_cases_deaths_spanish$Grupo_de_Edad, pattern=fixed('Age_'), replacement=fixed('') )

age_new_cases_deaths_spanish

write.csv(age_new_cases_deaths_spanish, file = "SpanishData/age_new_cases_deaths_spanish.csv")


```


#Rename Age_Master in Spanish 
```{r}
age_master_spanish <- age_master

age_master_spanish <- age_master_spanish %>% 
  rename(Casos_0_10 = Cases_0_10, Casos_11_17 = Cases_11_17, Casos_18_24 = Cases_18_24,Casos_25_34 = Cases_25_34, Casos_35_44 = Cases_35_44, Casos_45_54 = Cases_55_64, Casos_65 = Cases_65, Casos_Desconocidos = Cases_Unknown, Muertes_0_10 = Deaths_0_10, Muertes_11_17 = Deaths_11_17, Muertes_18_24 = Deaths_18_24, Muertes_25_34 = Deaths_25_34, Muertes_35_44 = Deaths_35_44, Muertes_45_54 = Deaths_45_54, Muertes_55_64 = Deaths_55_64, Muertes_65 = Deaths_65, Muertes_Desconocidos = Deaths_Unknown, Casos_0_17 = Cases_0_17, Casos_25_44 = Cases_25_44, Casos_45_64 = Cases_45_64, Nuevos_Casos_0_10 = New_Cases_0_10, Nuevos_Casos_11_17 = New_Cases_11_17, Nuevos_Casos_18_24 = New_Cases_18_24,Nuevos_Casos_25_34 = New_Cases_25_34, Nuevos_Casos_35_44 = New_Cases_35_44, Nuevos_Casos_45_54 = New_Cases_45_54, Nuevos_Casos_55_64 = New_Cases_55_64, Nuevos_Casos_65 = New_Cases_65, Nuevas_Muertes_0_10 = New_Deaths_0_10, Nuevas_Muertes_11_17 = New_Deaths_11_17, Nuevas_Muertes_18_24 = New_Deaths_18_24, Nuevas_Muertes_25_34 = New_Deaths_25_34, Nuevas_Muertes_35_44 = New_Deaths_35_44, Nuevas_Muertes_45_54 = New_Deaths_45_54,  Nuevas_Muertes_55_64 = New_Deaths_55_64, Nuevas_Muertes_65 = New_Deaths_65, Muertes_0_17 = Deaths_0_17, Muertes_25_44 = Deaths_25_44, Muertes_45_64 = Deaths_45_64, Nuevos_Casos_0_17 = New_Cases_0_17, Nuevos_Casos_25_44 = New_Cases_25_44, Nuevos_Casos_45_64 = New_Cases_45_64, Nuevas_Muertes_0_17 = New_Deaths_0_17, Nuevas_Muertes_25_44 = New_Deaths_25_44, Nuevas_Muertes_45_64 = New_Deaths_45_64) %>% 
  filter(Date >= "2020-09-13")

age_master_spanish$Date <- as.Date(age_master_spanish$Date)


#age_master_spanish <- janitor::clean_names(age_master_spanish)

glimpse(age_master_spanish)

write.csv(age_master_spanish, file = "SpanishData/age_FLOURISH_spanish_master.csv")
```

#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#RACE calculations
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#Today's data with race
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#RACE calculations
#----------------------------------------------------------------------------------------------------#
#----------------------------------------------------------------------------------------------------#
#Today's data with race
```{r}
today_race <- today_demo 
  
  
```  


```{r}
today_race <- today_race %>%
  rename(Date=Date, Cases_African_American=Black,
Cases_White=White,
Cases_Asian=Asian,
Cases_Native_American=Na,
Cases_Pacific_Islander=Pi,
Cases_Other=Other_Race,
Cases_Race_Or_Ethnicity_Unknown=Unk_Race,
Cases_Multi_Racial=Multi_Race,
Cases_Hispanic=Hispanic,
Deaths_African_American=Deaths_Black,
Deaths_White=Deaths_White,
Deaths_Asian=Deaths_Asian,
Deaths_Native_American=Deaths_Na,
Deaths_Pacific_Islander=Deaths_Pi,
Deaths_Other_Race=Deaths_Other_Race,
Deaths_Race_Or_Ethnicity_Unknown=Deaths_Unk_Race,
Deaths_Multi_Racial=Deaths_Multi_Race,
Deaths_Hispanic=Deaths_Hispanic)

glimpse(today_race)
```

#Today_yesterday
#Import the master file for race from GitHub

```{r}
race3 <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/race_main_file.csv")
#race3 <- rio::import("race_main_file.csv")
race3$Date <- lubridate::mdy(race3$Date)

glimpse(race3)
```


```{r}
race3 <- as.data.frame(race3)
yesterday_race <- race3 %>% 
  select("Date", "Cases_African_American", "Cases_White", "Cases_Native_American", "Cases_Asian", "Cases_Pacific_Islander", "Cases_Hispanic", "Cases_Multi_Racial", "Deaths_African_American", "Deaths_White", "Deaths_Native_American", "Deaths_Asian", "Deaths_Pacific_Islander", "Deaths_Hispanic", "Deaths_Multi_Racial", "Total_State_Population_African_American", "Total_State_Population_Asian", "Total_State_Population_Hispanic", "Total_State_Population_Multi_Racial", "Total_State_Population_Native_American",
"Total_State_Population_Pacific_Islander", "Total_State_Population_White",  "Total_State_Population_Other", "Total_State_Population_Race_Or_Ethnicity_Unknown", "Pct_State_Population_African_American",  "Pct_State_Population_Asian", "Pct_State_Population_Hispanic", "Pct_State_Population_Multi_Racial",  "Pct_State_Population_Native_American", "Pct_State_Population_Pacific_Islander", 
"Pct_State_Population_White", "Pct_State_Population_Other", "Pct_State_Population_Race_Or_Ethnicity_Unknown") %>% 
    filter(Date == yesterday)


#if you miss a day, fill this out and run instead
# filter(date == "2020-10-13") %>% 

yesterday_race$Date <- as.Date(yesterday_race$Date)
glimpse(yesterday_race)
```

#Create a population table
```{r}
pop <- race3 %>% 
  select("Date", "Total_State_Population_African_American", "Total_State_Population_Asian", "Total_State_Population_Hispanic", "Total_State_Population_Multi_Racial", "Total_State_Population_Native_American",
"Total_State_Population_Pacific_Islander", "Total_State_Population_White",  "Total_State_Population_Other", "Total_State_Population_Race_Or_Ethnicity_Unknown", "Pct_State_Population_African_American",  "Pct_State_Population_Asian", "Pct_State_Population_Hispanic", "Pct_State_Population_Multi_Racial",  "Pct_State_Population_Native_American", "Pct_State_Population_Pacific_Islander", 
"Pct_State_Population_White", "Pct_State_Population_Other", "Pct_State_Population_Race_Or_Ethnicity_Unknown") %>% 
  filter(Date == yesterday) 

#if you miss a day, fill this out and run instead
#filter(date == "2020-10-13") %>% 
#
#Change to today's date
pop$Date <- today
#if you miss a day, fill this out and run instead
#pop$date <- "2020-10-14"

pop$Date <- as.Date(pop$Date)
glimpse(pop)
```


```{r}
today_race2 <- today_race %>% 
   left_join(pop, by="Date")
head(today_race2)
```

#You should get a df with 2 observations (two days' worth a data)
```{r}
twodays2 <- smartbind(today_race2, yesterday_race)
twodays2$Date <- as.Date(twodays2$Date)
glimpse(twodays2)

```

-**The Today-Yesterday Calculations**

```{r}
twodays3 <- twodays2 %>%
  mutate(New_Cases_African_American = (Cases_African_American-lead(Cases_African_American))) %>% 
  mutate(New_Cases_White = (Cases_White-lead(Cases_White))) %>% 
   mutate(New_Cases_Native_American = (Cases_Native_American-lead(Cases_Native_American))) %>%
   mutate(New_Cases_Asian = (Cases_Asian-lead(Cases_Asian))) %>%
   mutate(New_Cases_Pacific_Islander = (Cases_Pacific_Islander-lead(Cases_Pacific_Islander))) %>%
   mutate(New_Cases_Hispanic = (Cases_Hispanic-lead(Cases_Hispanic))) %>%
   mutate(New_Cases_Multi_Racial = (Cases_Multi_Racial-lead(Cases_Multi_Racial))) %>%
   mutate(New_Deaths_African_American = (Deaths_African_American-lead(Deaths_African_American))) %>%
   mutate(New_Deaths_White = (Deaths_White-lead(Deaths_White))) %>%
   mutate(New_Deaths_Native_American = (Deaths_Native_American-lead(Deaths_Native_American))) %>%
   mutate(New_Deaths_Asian = (Deaths_Asian-lead(Deaths_Asian))) %>%
   mutate(New_Deaths_Pacific_Islander = (Deaths_Pacific_Islander-lead(Deaths_Pacific_Islander))) %>%
   mutate(New_Deaths_Hispanic = (Deaths_Hispanic-lead(Deaths_Hispanic))) %>%
   mutate(New_Deaths_Multi_Racial = (Deaths_Multi_Racial-lead(Deaths_Multi_Racial))) 
```

#Percentage Calculations 
```{r}
twodays3 <- twodays3 %>%
mutate(Cases_Per_10_000_African_American = (Cases_African_American / Total_State_Population_African_American)*10000) %>% 
mutate(Cases_Per_10_000_Asian = (Cases_Asian / Total_State_Population_Asian)*10000) %>% 
mutate(Cases_Per_10_000_Multi_Racial = (Cases_Multi_Racial / Total_State_Population_Multi_Racial)*10000) %>%
mutate(Cases_Per_10_000_Pacific_Islander = (Cases_Pacific_Islander / Total_State_Population_Pacific_Islander)*10000) %>%
mutate(Cases_Per_10_000_Hispanic = (Cases_Hispanic / Total_State_Population_Hispanic)*10000) %>%
mutate(Cases_Per_10_000_Native_American = (Cases_Native_American / Total_State_Population_Native_American)*10000) %>%
mutate(Cases_Per_10_000_White = (Cases_White / Total_State_Population_White)*10000) %>%
mutate(Deaths_Per_10_000_African_American = (Deaths_African_American / Total_State_Population_African_American)*10000) %>% 
mutate(Deaths_Per_10_000_Asian = (Deaths_Asian / Total_State_Population_Asian)*10000) %>% 
mutate(Deaths_Per_10_000_Multi_Racial = (Deaths_Multi_Racial / Total_State_Population_Multi_Racial)*10000) %>%
mutate(Deaths_Per_10_000_Pacific_Islander = (Deaths_Pacific_Islander / Total_State_Population_Pacific_Islander)*10000) %>%
mutate(Deaths_Per_10_000_Hispanic = (Deaths_Hispanic / Total_State_Population_Hispanic)*10000) %>%
mutate(Deaths_Per_10_000_Native_American = (Deaths_Native_American / Total_State_Population_Native_American)*10000) %>%
mutate(Deaths_Per_10_000_White = (Deaths_White / Total_State_Population_White)*10000)

```


#Merges the Today's calculations with the master file
```{r}
today_race_calcs <- twodays3 %>% 
  filter(Date ==today)
#if you miss a day, fill this out and run instead
#filter(date == "2020-10-14")
  
  
#today_race_calcs <- janitor::clean_names(today_race_calcs)
glimpse(today_race_calcs)
```

#Writing the Final Updated File


```{r}
race_main_file <- smartbind(race3, today_race_calcs)
#Cut rows from df
#race_main_file <- slice(race_main_file, -c(1))

#Cut Columns
race_main_file <- race_main_file[ -c(1) ]

race_main_file <- race_main_file %>% 
  arrange(desc(Date))

write.csv(race_main_file, file = "MasterData/race_main_file.csv")
write.csv(race_main_file, file = "Demographics/race_main_file.csv")
```



#--------------------------------
#New Cases Today Table for Flourish
#--------------------------------
#Rename variables here

```{r}
write.csv(race_main_file, file = "Demographics/race_FLOURISH_main_file.csv")
```

#Race New Cases for Flourish
```{r}
x <- race_main_file %>% 
  select(Date, New_Cases_African_American, New_Cases_Asian, New_Cases_Hispanic, New_Cases_Multi_Racial, New_Cases_Native_American, New_Cases_Other, New_Cases_Race_Or_Ethnicity_Unknown, New_Cases_White) %>% 
  filter(Date == today) %>% 
  distinct()

#flip
race_new_cases <- gather(x, key = "Race", value = "New_Cases", New_Cases_African_American, New_Cases_Asian, New_Cases_Hispanic, New_Cases_Multi_Racial, New_Cases_Native_American, New_Cases_Other, New_Cases_Race_Or_Ethnicity_Unknown, New_Cases_White)

write.csv(race_new_cases, file = "Demographics/race_new_cases.csv")
race_new_cases
```


#Spanish translation race_new_cases

#Renaming for Spanish
```{r}
race_main_file_spanish <- race_main_file %>% 
  rename(Fecha=Date, Casos_Afroamericano = Cases_African_American,
Casos_Blanco = Cases_White,
Casos_Asiático = Cases_Asian,
Casos_Nativo_Americano = Cases_Native_American,
Casos_Oriundo_de_las_Islas_del_Océano_Pacífico = Cases_Pacific_Islander,
Casos_Multirracial = Cases_Multi_Racial,
Casos_Hispano = Cases_Hispanic,
Muertes_Afroamericanos = Deaths_African_American,
Muertes_Blanco = Deaths_White,
Muertes_Asiático = Deaths_Asian,
Muertes_Nativo_Americano = Deaths_Native_American,
Muertes_Oriundo_de_las_Islas_del_Océano_Pacífico = Deaths_Pacific_Islander,
Muertes_Multirracial = Deaths_Multi_Racial,
Muertes_Hispano = Deaths_Hispanic,
Nuevos_Casos_Afroamericano = New_Cases_African_American,
Nuevos_Casos_Blanco = New_Cases_White,
Nuevos_Casos_Asiático = New_Cases_Asian,
Nuevos_Casos_Nativo_Americano = New_Cases_Native_American,
Nuevos_Casos_Oriundo_de_las_Islas_del_Océano_Pacífico = New_Cases_Pacific_Islander,
Nuevos_Casos_Multirracial = New_Cases_Multi_Racial,
Nuevos_Casos_Hispano = New_Cases_Hispanic,
Nuevas_Muertes_Afroamericano = New_Deaths_African_American,
Nuevas_Muertes_Blanco = New_Deaths_White,
Nuevas_Muertes_Asiático = New_Deaths_Asian,
Nuevas_Muertes_Nativo_Americano = New_Deaths_Native_American,
Nuevas_Muertes_Oriundo_de_las_Islas_del_Océano_Pacífico = New_Deaths_Pacific_Islander,
Nuevas_Muertes_Multirracial = New_Deaths_Multi_Racial,
Nuevas_Muertes_Hispano = New_Deaths_Hispanic,
Casos_por_cada_10k_Afroamericano = Cases_Per_10_000_African_American,
Casos_por_cada_10k_Blanco = Cases_Per_10_000_White,
Casos__por_cada_10k_Asiático = Cases_Per_10_000_Asian,
Casos_por_cada_10k_Nativo_Americano = Cases_Per_10_000_Native_American,
Casos_por_cada_10k_Oriundo_de_las_Islas_del_Océano_Pacífico = Cases_Per_10_000_Pacific_Islander,
Casos_por_cada_10k_Multirracial = Cases_Per_10_000_Multi_Racial,
Casos_por_cada_10k_Hispano = Cases_Per_10_000_Hispanic,
Muertes_por_cada_10k_Afroamericanos = Deaths_Per_10_000_African_American,
Muertes_por_cada_10k_Blanco = Deaths_Per_10_000_White,
Muertes_por_cada_10k_Asiático = Deaths_Per_10_000_Asian,
Muertes_por_cada_10k_Nativo_Americano = Deaths_Per_10_000_Native_American,
Muertes_por_cada_10k_Oriundo_de_las_Islas_del_Océano_Pacífico = Deaths_Per_10_000_Pacific_Islander,
Muertes_por_cada_10k_Multirracial = Deaths_Per_10_000_Multi_Racial,
Muertes_por_cada_10k_Hispano = Deaths_Per_10_000_Hispanic)

race_main_file_spanish
write.csv(race_main_file_spanish, file = "SpanishData/race_main_file_spanish.csv")

```


#Race New Cases in Spanish
```{r}
x <- race_main_file_spanish %>% 
  select(Fecha, Nuevos_Casos_Afroamericano, Nuevos_Casos_Asiático, Nuevos_Casos_Hispano, Nuevos_Casos_Multirracial, Nuevos_Casos_Nativo_Americano, Nuevos_Casos_Blanco) %>% 
  filter(Fecha == today) %>% 
  distinct()

#flip
race_new_cases_spanish <- gather(x, key = "Razo o Ethnicidad", value = "Nuevos_Casos", Nuevos_Casos_Afroamericano, Nuevos_Casos_Asiático, Nuevos_Casos_Hispano, Nuevos_Casos_Multirracial, Nuevos_Casos_Nativo_Americano, Nuevos_Casos_Blanco)


write.csv(race_new_cases_spanish, file = "SpanishData/race_new_cases_spanish.csv")
race_new_cases_spanish
```


#New Deaths Today Table for Flourish
```{r}
race_new_deaths <- race_main_file %>% 
  select(Date, New_Deaths_African_American, New_Deaths_Asian, New_Deaths_Hispanic, New_Deaths_Multi_Racial, New_Deaths_Native_American, New_Deaths_Other, New_Deaths_Race_Or_Ethnicity_Unknown, New_Deaths_White) %>% 
  filter(Date == today) %>% 
  distinct()

#flip
race_new_deaths <- gather(race_new_deaths, key = "Race", value = "New_Deaths", New_Deaths_African_American, New_Deaths_Asian, New_Deaths_Hispanic, New_Deaths_Multi_Racial, New_Deaths_Native_American, New_Deaths_Other, New_Deaths_Race_Or_Ethnicity_Unknown, New_Deaths_White)

race_new_deaths
write.csv(race_new_deaths, file = "Demographics/race_new_deaths.csv")
```

#Race New Deaths in Spanish
```{r}
race_new_deaths_spanish <- race_main_file_spanish %>% 
  select(Fecha, Nuevas_Muertes_Afroamericano, Nuevas_Muertes_Asiático, Nuevas_Muertes_Hispano, Nuevas_Muertes_Multirracial, Nuevas_Muertes_Nativo_Americano, Nuevas_Muertes_Blanco) %>% 
  filter(Fecha == today) %>% 
  distinct()

#flip
race_new_deaths_spanish <- gather(race_new_deaths_spanish, key = "Razo o Ethnicidad", value = "Nuevas_Muertes", Nuevas_Muertes_Afroamericano, Nuevas_Muertes_Asiático, Nuevas_Muertes_Hispano, Nuevas_Muertes_Multirracial, Nuevas_Muertes_Nativo_Americano, Nuevas_Muertes_Blanco)


write.csv(race_new_deaths_spanish, file = "SpanishData/race_new_deaths.csv")
race_new_deaths_spanish

```

#COMBINE race_new_cases and race new deaths

```{r}
head(race_new_cases)
```


```{r}
#SF$disposition1 <- str_replace_all(SF$disposition1, pattern=fixed('ABA'), replacement=fixed('Abated') )
#strip the race name down
race_new_cases$NEW <- str_replace_all(race_new_cases$Race, pattern=fixed('New_Cases_'), replacement=fixed(''))
race_new_deaths$NEW <- str_replace_all(race_new_deaths$Race, pattern=fixed('New_Deaths_'), replacement=fixed(''))

```

```{r}
race_new_cases_deaths <- race_new_cases %>% 
   inner_join(race_new_deaths, by=c("NEW", "Date"))

race_new_cases_deaths <- race_new_cases_deaths %>% 
  select(Date, NEW, New_Cases, New_Deaths)

colnames(race_new_cases_deaths)[2] <- "Race_Ethnicity"

write.csv(race_new_cases_deaths, file = "Demographics/race_new_cases_deaths.csv")
head(race_new_cases_deaths)

#Spanish translation

#Spanish translation

race_new_cases_deaths_spanish <- race_new_cases_deaths %>% 
  rename(Nuevas_Muertes = New_Deaths, Fecha = Date, Nuevos_Casos = New_Cases, Razo_o_Ethnicidad=Race_Ethnicity) 

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('African_American'), replacement=fixed('Afroamericano') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('Asian'), replacement=fixed('Asiático') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('Hispanic'), replacement=fixed('Hispano') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('Multi_Racial'), replacement=fixed('Multirracial') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('Native_American'), replacement=fixed('Nativo_Americano') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('Pacific_Islander'), replacement=fixed('de_las_Islas_del_Océano_Pacífico') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('Other'), replacement=fixed('Otro') )

race_new_cases_deaths_spanish$ Razo_o_Ethnicidad <- str_replace_all(race_new_cases_deaths_spanish$ Razo_o_Ethnicidad, pattern=fixed('White'), replacement=fixed('Blanco') )


race_new_cases_deaths_spanish

write.csv(race_new_cases_deaths_spanish, file = "SpanishData/race_new_cases_deaths_spanish.csv")


```

-----------------------------------------------------------------------#
#Gender
#----------------------------------------------------------------------------------------------------#
#Today's data 
```{r}
today_gender <- today_demo %>% 
  select(Date, Male, Female, Na)
```

#Rename

```{r}
today_gender2 <- today_gender

today_gender2 <- today_gender2 %>% 
  rename(Cases_Male = Male, Cases_Female = Female, Cases_Gender_Unknown = Na)

today_gender2$Date <- as.Date(today_gender2$Date)

#today_gender2 <- janitor::clean_names(today_gender2)

glimpse(today_gender2)
```

#Today_yesterday
#Import the master file for gender

```{r}
gender_master <- rio::import("gender_master.csv")
#gender_master <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/gender_master.csv")
gender_master$Date <- lubridate::ymd(gender_master$Date)


head(gender_master)
```

```{r}
yesterday_gender <- gender_master %>% 
   filter(Date == yesterday)

yesterday_gender <- gender_master %>% 
  select(Date, Cases_Male, Cases_Female, Cases_Gender_Unknown)

glimpse(yesterday_gender)
```

#You should get a df with 2 observations (two days' worth a data)
```{r}
twodays3 <- smartbind(today_gender2, yesterday_gender)
twodays3$Date <- as.Date(twodays3$Date)
# twodays3 <- twodays3 %>% 
#   distinct()
#cut an extra day off
#twodays3 <- slice(twodays3, -c(1))
glimpse(twodays3)

```

-**The Today-Yesterday Calculations**

```{r}
gender2 <- twodays3 %>%
mutate(New_Cases_Male = (Cases_Male -lead(Cases_Male))) %>% 
  mutate(New_Cases_Female = (Cases_Female-lead(Cases_Female))) %>% 
  mutate(New_Cases_Gender_Unknown = (Cases_Gender_Unknown-lead(Cases_Gender_Unknown)))

head(gender2)

write.csv(gender2, file = "Demographics/gender_master.csv")
write.csv(gender2, file = "MasterData/gender_master.csv")
write.csv(gender2, file = "Demographics/gender_FLOURISH_master.csv")
```

#New Cases Today Table for Flourish


```{r}
gender_new_cases <- gender2 %>% 
  select(Date, "New_Cases_Male", "New_Cases_Female","New_Cases_Gender_Unknown") %>% 
  filter(Date == today)

#flip
gender_new_cases <- gather(gender_new_cases, key = "Gender", value = "New_Cases", "New_Cases_Male", "New_Cases_Female","New_Cases_Gender_Unknown")

gender_new_cases
write.csv(gender_new_cases, file = "Demographics/gender_new_cases.csv")

#Spanish Translation

write.csv(gender_new_cases_spanish, file = "SpanishData/gender_new_cases_spanish.csv")
gender_new_cases_spanish

```


#Gender_Master in Spanish
```{r}
gender_spanish_master <- gender2 %>% 
  rename(Fecha = Date, Casos_Hombres = Cases_Male, Casos_Mujeres = Cases_Female, Casos_Género_Desconocidos = Cases_Gender_Unknown, Nuevos_Casos_Hombres = New_Cases_Male, Nuevos_Casos_Mujeres = New_Cases_Female, Nuevos_Casos_Género_Desconocidos = New_Cases_Gender_Unknown)

write.csv(gender_spanish_master, file = "SpanishData/gender_FLOURISH_master_spanish.csv")

```

#More Renaming for Spanish
```{r}
gender_new_cases_spanish <- gender_spanish_master %>% 
  select(Fecha, "Nuevos_Casos_Hombres", "Nuevos_Casos_Mujeres","Nuevos_Casos_Género_Desconocidos") %>% 
  filter(Fecha == today)

#flip
gender_new_cases_spanish <- gather(gender_new_cases_spanish, key = "Género", value = "Nuevos_Casos", "Nuevos_Casos_Hombres", "Nuevos_Casos_Mujeres","Nuevos_Casos_Género_Desconocidos")


write.csv(gender_new_cases_spanish, file = "SpanishData/gender_new_cases_spanish.csv")
gender_new_cases_spanish
```




-**30**


