---
title: "DIRECT API PULL Arkansas Covid Daily Update - County"
author: "Rob Wells"
date: "8/28/2020"
output: html_document
---

# Retrieve Arkansas Covid Data from ADH 


```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(tidyr)
library(httr)
library(jsonlite)

#install.packages("googledrive")
#install.packages("googlesheets4")
#library(googlesheets4)
#library("googledrive")
# vignette("basic-usage", package="googlesheets")
```


This Works. 
Question: Missing a Date Field.  
And here is the daily data
```{r}
#Austin's query from json feed
q <- fromJSON('https://services.arcgis.com/PwY9ZuZRDiI5nXUB/ArcGIS/rest/services/ADH_COVID19_Positive_Test_Results/FeatureServer/0/query?where=0%3D0&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&resultType=none&distance=0.0&units=esriSRUnit_Meter&returnGeodetic=false&outFields=county_nam%2C+positive%2C+negative%2C+Recoveries%2C+deaths%2C+total_tests%2C+active_cases&returnGeometry=false&returnCentroid=false&featureEncoding=esriDefault&multipatchOption=xyFootprint&maxAllowableOffset=&geometryPrecision=&outSR=&datumTransformation=&applyVCSProjection=false&returnIdsOnly=false&returnUniqueIdsOnly=false&returnCountOnly=false&returnExtentOnly=false&returnQueryGeometry=false&returnDistinctValues=false&cacheHint=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&having=&resultOffset=&resultRecordCount=&returnZ=false&returnM=false&returnExceededLimitFeatures=true&quantizationParameters=&sqlFormat=none&f=pjson&token=')

api_pull <- q[["features"]][["attributes"]]
#print's today's date as a date field
#api_pull$date <- Sys.Date()
#file.exists(filename)
#austin help on that
#if file exists = check if current data is different and if that is true then append today's data set to master dataset and resave
rm(q)
```

```{r}
api_pull <- api_pull %>% 
  arrange(county_nam)
head(api_pull)
```

# Part 1: Fix the data structure

- **Clean Names, Rename**


```{r}

api_pull <- janitor::clean_names(api_pull)

  #rename(NEW NAME = EXISTING NAME) %>% 
api_pull <- api_pull %>% 
  rename(county = county_nam) %>% 
  rename(recovered = recoveries) %>%
  rename(total_deaths = deaths) %>%
  rename(current_infections = total_tests) 
```

- **Adding Date Field**
```{r}
api_pull <- api_pull %>% 
  mutate(date = Sys.Date()) 

```



- **Each Day:** We join the new data, api, with the population and then perform the calculations

```{r}
population <- rio::import("./Exercises/population.csv")
api_pull <- api_pull %>% 
  inner_join(population, by="county")
rm(population)
head(api_pull)

```
**This is the update checker **
Check if covid_master.csv does NOT exists and create one,
otherwise load covid_master.csv as master, and change factor variables to character&date
Then check if the last 75 enteries are identical in columns 1:5,
if so, data is not updated,
else, bind new api_pull to master and save

Either way, it outputs a variable "hold", which is the most current dataframe pulled from the api
```{r}
if (!file.exists("Data/aoi_hold.csv")) {
  write.csv(api_pull,"Data/api_hold.csv",row.names = FALSE)
  hold <- api_pull
} else {
  hold <- read.csv("Data/api_hold.csv")
  hold[1] <- lapply(hold[1], as.character)
  hold[8] <- lapply(hold[8], as.Date)
  if (identical(api_pull[1:75,1:5],hold[1:75,1:5])){
    print("not updated")
    quit(save = "no")
  } else{
    hold <- bind_rows(api_pull,hold)
    write.csv(hold,"Data/api_hold.csv",row.names = FALSE)
  }
}

```

# Part 2: MATH!   

```{r}
master <- hold %>%
  group_by(county) %>% 
  mutate(new_cases_today = (positive-lead(positive))) %>%
  mutate(recovered_since_yesterday = (recovered-lead(recovered))) %>%
  mutate(new_deaths_today = (total_deaths-lead(total_deaths))) %>%
  mutate(number_tested = (positive + negative)) %>% 
  mutate(new_tests_dashboard = (number_tested-lead(number_tested)))

```



-**The Percentage Calculations**

This calculates
Column E (Cases/population = (positive / population) *100
Column F (Tested/population = (Number tested / population) *100
Column G (New Cases Today per 10k Population = (New Cases Today / population) *10000
Column H (Active Cases per 10k = (current_infections / population) *10000
Column T (% Positive Cumulative = (positive / Number tested) *100
Column U (% Positive New to Dashboard = New Cases Today / New Tests Dashboard) * 100
Column V (Closed = Recovered + Total Deaths)
Column W (% Deaths vs. Recoveries = Total Deaths / Closed) * 100
Column X (% Recoveries vs. Deaths = Recovered / Closed) * 100


```{r}
master <- master %>%
  mutate(cases_population = (positive / population)*100) %>% 
  mutate(tested_population = (number_tested / population)*100) %>% 
  mutate(new_cases_today_10k_pop = (new_cases_today/population)*10000) %>% 
  mutate(active_cases_10k_pop = (current_infections/population)*10000) %>% 
  mutate(pct_positive_cumulative = (positive/number_tested)*100) %>% 
  mutate(pct_positive_new_to_dashboard = (new_cases_today/new_tests_dashboard)*100) %>% 
  mutate(closed = (recovered + total_deaths)) %>% 
  mutate(pct_deaths_vs_recoveries = (total_deaths/closed)*100) %>% 
  mutate(pct_recoveries_vs_deaths = (recovered/closed)*100)
  

```


align names in order of covid table
```{r}
master <- master %>% 
  select(county,date,population,cases_population,tested_population,new_cases_today_10k_pop,active_cases_10k_pop,current_infections,positive, new_cases_today, negative, recovered, recovered_since_yesterday, total_deaths, new_deaths_today, number_tested, new_tests_dashboard, pct_positive_cumulative, pct_positive_new_to_dashboard, closed, pct_deaths_vs_recoveries, pct_recoveries_vs_deaths)

```

colnames(covid_temp)[6,7,18,19,21,22] <- c("new_cases_today_10k_pop", "active_cases_10k_pop", "pct_positive_cumulative", "pct_positive_new_to_dashboard", "pct_deaths_vs_recoveries", "pct_recoveries_vs_deaths")



RENAME COLUMNS TO MATCH NEW FILE
```{r}
#I don't know which names are correct, but this throws an error, so I'm commenting it out for now
# -Austin
# master2 <- master2 %>% 
#   #rename(NEW NAME = EXISTING NAME) %>% 
#   rename(new_cases_today_10k_pop = new_cases_today_per_10k_population) %>% 
#   rename(active_cases_10k_pop = active_cases_per_10k) %>% 
#   rename(pct_positive_cumulative = percent_positive_cumulative) %>% 
#   rename(pct_positive_new_to_dashboard = percent_positive_new_to_dashboard) %>% 
#   rename(pct_deaths_vs_recoveries = percent_deaths_vs_recoveries) %>% 
#   rename(pct_recoveries_vs_deaths = percent_recoveries_vs_deaths)
```    


**Final **
The variable
```{r}

write_csv(master, "Data/covid_master.csv")
```
