---
title: "School Data"
author: "Rob Wells and Katy Seiter" 
date: "11/24/2020"
output: pdf_document
---

# Compilation of School COVID-19 Reports for Arkansascovid.com Calculations

- **This notebook retrieves data from a Google sheet and compiles a time series of ADH school infection reports**

--**This 11/24 version created a time series from Aug 27-Nov 19**

-**It creates a masterfile: schools_master.csv**

# Part 1: Import Data, Clean It

```{r include=FALSE}
#install.packages("slider")
#install.packages("zoo")
#install.packages("gtools")
# vignette("basic-usage", package="googlesheets")
#install.packages("googledrive")
#install.packages("googlesheets4")
library(tidyverse)
library(janitor)
library(lubridate)
library(googlesheets4)
library(googledrive)
library(tidyr)
library(jsonlite)
library(gtools)
library(zoo)
library(reshape2)
library(slider)
```
#This section for importing the master table and creating the master file just once
#For Googlesheet access
```{r include=FALSE}
url2 <- ("https://docs.google.com/spreadsheets/d/1UmaSSoMbou2MpvGmihP5cZPNqBnaxqIH60cs10sShB8/edit#gid=748934096")

school12_3 <- read_sheet(url2, sheet = "master",  col_types = NULL, na = "",trim_ws = TRUE)

colnames(school12_3)[1] <-c("date")
```

```{r}
school12_3 <- janitor::clean_names(school12_3)
glimpse(school12_3)
```
#Subset tables
#Public School

```{r}
#Rename
school12_3_bak <- school12_3
school12_3 <- as.data.frame(school12_3)
#as.numeric(unlist(school12_3$ps_cumulative_faculty_staff_cases))
#SF$disposition1 <- str_replace_all(SF$disposition1, pattern=fixed('ABA'), replacement=fixed('Abated') )

school12_3$ps_cumulative_faculty_staff_cases <- str_replace_all(school12_3$ps_cumulative_faculty_staff_cases, 
                                                                 pattern=fixed('NULL'), replacement=fixed('0'))
school12_3$ps_cumulative_student_cases  <- str_replace_all(school12_3$ps_cumulative_student_cases, 
                                                                 pattern=fixed('NULL'), replacement=fixed('0'))

school12_3[4:5] <- lapply(school12_3[4:5], as.numeric)
#glimpse(school12_3)



#Subset public schools
 ps <- school12_3 %>% 
   select("date", "public_school_district","ps_active_cases", "ps_cumulative_faculty_staff_cases", "ps_cumulative_student_cases", "ps_cumulative_total")
 
#create type
 ps <- ps %>%
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total")

#ps <- as.data.frame(ps)
ps
```

#Private School

```{r}
#Subset private schools
 private <- school12_3 %>% 
   select("date", "private_school","private_active_cases", "private_cumulative_faculty_staff_cases",                   "private_cumulative_student_cases", "private_cumulative_total")
#create type
private <- private %>% 
   mutate(type="private_schools")

#rename columns
colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

private <- private %>% 
  distinct()
private <- as.data.frame(private)
private
```

#College

```{r}
#Subset private schools
 college <- school12_3 %>% 
   select("date", "college_university","uni_active_cases", "uni_cumulative_faculty_staff_cases",                   "uni_cumulative_student_cases", "uni_cumulative_total")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school12_3)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school12_3 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","x5_cumulative_active_cases", "x5_cumulative_faculty_staff_cases", "x5_cumulative_student_cases", "x5_cumulative_total")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
[22] "totals_for_all_schools"                            
[23] "cumulative_active_cases"                           
[24] "cumulative_faculty_staff_cases"                    
[25] "cumulative_student_cases"                          
[26] "cumulative_total"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school12_3 %>% 
   select("date", "totals_for_all_schools","cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
schools_master8 <- smartbind(ps,private,college,less5,totals_all_schools)
head(schools_master8)
write.csv(schools_master8, "schools_master.csv")
# masterschools <- rio::import("https://raw.githubusercontent.com/profrobwells/CovidFall2020/master/schools_master.csv")
# schools_master <- smartbind(schools_master8, masterschools)
#write.csv(schools_master, "schools_master.csv")
```





#This section for building the table from individual sheets

#For Googlesheet access
```{r include=FALSE}
url2 <- ("https://docs.google.com/spreadsheets/d/1UmaSSoMbou2MpvGmihP5cZPNqBnaxqIH60cs10sShB8/edit#gid=748934096")

school12_3 <- read_sheet(url2, sheet = "Nov19",  col_types = NULL, na = "",trim_ws = TRUE)
school12_3$date <- "2020-11-19"
```


```{r include=FALSE}
# school12_3 <- rio::import("SchoolMaster.xlsx", which = "Nov19")
# school10_15 <- rio::import("SchoolMaster.xlsx", which = "Oct15")
# school10_15 <- rio::import("SchoolMaster.xlsx", which = "Oct15")
# school10_15$date <- "2020-10-15"
# school10_12 <- rio::import("SchoolMaster.xlsx", which = "Oct12")
# school10_8 <- rio::import("SchoolMaster.xlsx", which = "Oct8")
# school10_5 <- rio::import("SchoolMaster.xlsx", which = "Oct5")
# school10_1 <- rio::import("SchoolMaster.xlsx", which = "Oct1")
# school9_28 <- rio::import("SchoolMaster.xlsx", which = "XXXSept28")
# school10_12$date <- "2020-10-12"
# school10_8$date <- "2020-10-08"
# school10_5$date <- "2020-10-05"
# school10_1$date <- "2020-10-01"
# school9_28$date <- "2020-09-28"
```


```{r}
#glimpse(school12_3)
```

```{r}
school12_3 <- janitor::clean_names(school12_3)
names(school12_3)
```

#Subset tables
#Public School

```{r}
#Rename
school12_3_bak <- school12_3
#Subset public schools
 ps <- school12_3 %>%
   select("date", "public_school_district","active_cases_2", "cumulative_faculty_staff_cases_3", "cumulative_student_cases_4", "cumulative_total_5")
 
 # ps <- school12_3 %>% 
 #   select("date", "public_school_district","active_cases_2", "cumulative_faculty_staff_cases_3", "cumulative_student_cases_4", "cumulative_total_5")

#create type
 ps <- ps %>% 
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

ps <- as.data.frame(ps)

```

#Private School

```{r}
#Subset private schools
 private <- school12_3 %>% 
   select("date", "private_school","active_cases_7", "cumulative_faculty_staff_cases_8",                   "cumulative_student_cases_9", "cumulative_total_10")
#create type
private <- private %>% 
   mutate(type="private_schools")

#rename columns
colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

private <- private %>% 
  distinct()
private <- as.data.frame(private)
private
```

#College

```{r}
#Subset private schools
 college <- school12_3 %>% 
   select("date", "college_university","active_cases_12", "cumulative_faculty_staff_cases_13",                   "cumulative_student_cases_14", "cumulative_total_15")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school12_3)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school12_3 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","cumulative_active_cases_17", "cumulative_faculty_staff_cases_18", "cumulative_student_cases_19", "cumulative_total_20")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
"totals_for_all_schools"                            
[22] "cumulative_active_cases_22"                        
[23] "cumulative_faculty_staff_cases_23"                 
[24] "cumulative_student_cases_24"                       
[25] "cumulative_total_25"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school12_3 %>% 
   select("date", "totals_for_all_schools","cumulative_active_cases_22", "cumulative_faculty_staff_cases_23", "cumulative_student_cases_24", "cumulative_total_25")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
schools_master <- smartbind(ps,private,college,less5,totals_all_schools)
head(schools_master)
masterschools <- rio::import("https://raw.githubusercontent.com/profrobwells/CovidFall2020/master/schools_master.csv")
schools_master <- smartbind(schools_master, masterschools)
write.csv(schools_master, "schools_master.csv")
```

#repeat lines 61-174 for each:
school10_12
school10_8
school10_5
school10_1

```{r}
school10_12 <- janitor::clean_names(school10_12)
names(school10_12)
```

#Subset tables
#Public School

```{r}
#Rename
school10_12_bak <- school10_12
#Subset public schools
 ps <- school10_12 %>% 
   select("date", "public_school_district","active_cases_2", "cumulative_faculty_staff_cases_3",                   "cumulative_student_cases_4", "cumulative_total_5")
#create type
 ps <- ps %>% 
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

ps <- as.data.frame(ps)

```

#Private School

```{r}
#Subset private schools
 private <- school10_12 %>% 
   select("date", "private_school","active_cases_7", "cumulative_faculty_staff_cases_8",                   "cumulative_student_cases_9", "cumulative_total_10")
#create type
private <- private %>% 
   mutate(type="private_schools")

#rename columns
colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

private <- private %>% 
  distinct()
private <- as.data.frame(private)
private
```

#College

```{r}
#Subset private schools
 college <- school10_12 %>% 
   select("date", "college_university","active_cases_12", "cumulative_faculty_staff_cases_13",                   "cumulative_student_cases_14", "cumulative_total_15")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school10_12)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school10_12 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","cumulative_active_cases_17", "cumulative_faculty_staff_cases_18", "cumulative_student_cases_19", "cumulative_total_20")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
"totals_for_all_schools"                            
[22] "cumulative_active_cases_22"                        
[23] "cumulative_faculty_staff_cases_23"                 
[24] "cumulative_student_cases_24"                       
[25] "cumulative_total_25"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school10_12 %>% 
   select("date", "totals_for_all_schools","cumulative_active_cases_22", "cumulative_faculty_staff_cases_23", "cumulative_student_cases_24", "cumulative_total_25")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
schools_master15 <- rio::import("https://raw.githubusercontent.com/profrobwells/CovidFall2020/master/schools_master.csv")
schools_master12 <- smartbind(ps,private,college,less5,totals_all_schools)
head(schools_master)
schools_master <- smartbind(schools_master12, schools_master15)
```

#school10_08
```{r}
school10_8 <- janitor::clean_names(school10_8)
names(school10_8)
```

#Subset tables
#Public School

```{r}
#Rename
school10_8_bak <- school10_8
#Subset public schools
 ps <- school10_8 %>% 
   select("date", "public_s_d_rict","ps_active_cases", "ps_cumulative_faculty_staff_cases", "ps_cumulative_student_cases", "ps_cumulative_total")
#create type
 ps <- ps %>% 
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

ps <- as.data.frame(ps)

```

#Private School

```{r}
#Subset private schools
 private <- school10_8 %>% 
   select("date", "private_school","pv_active_cases", "pv_cumulative_faculty_staff_cases",                   "pv_cumulative_student_cases", "pv_cumulative_total")
#create type
private <- private %>% 
   mutate(type="private_schools")

#rename columns
colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

private <- private %>% 
  distinct()
private <- as.data.frame(private)
private
```

#College

```{r}
#Subset private schools
 college <- school10_8 %>% 
   select("date", "college_university","uni_active_cases", "uni_cumulative_faculty_staff_cases",                   "uni_cumulative_student_cases", "uni_cumulative_total")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school10_8)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school10_8 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","cumulative_active_cases_less_than_5", "cumulative_faculty_staff_cases_less_than_5", "cumulative_student_cases_less_than_5", "cumulative_total_less_than_5")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
"totals_for_all_schools"                            
[22] "cumulative_active_cases_22"                        
[23] "cumulative_faculty_staff_cases_23"                 
[24] "cumulative_student_cases_24"                       
[25] "cumulative_total_25"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school10_8 %>% 
   select("date", "totals_for_all_schools","all_schools_cumulative_active_cases", "all_schools_cumulative_faculty_staff_cases", "all_schools_cumulative_student_cases", "all_schools_cumulative_total")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
schools_master8 <- smartbind(ps,private,college,less5,totals_all_schools)
head(schools_master8)
schools_master <- smartbind(schools_master8, schools_master)
```

#school10_05

```{r}
school10_5 <- janitor::clean_names(school10_5)
names(school10_5)
```

#Subset tables
#Public School

```{r}
#Rename
school10_05_bak <- school10_5
#Subset public schools
 ps <- school10_5 %>% 
   select("date", "public_school_district","ps_active_cases", "ps_cumulative_faculty_staff_cases",                   "ps_cumulative_student_cases", "ps_cumulative_total")
#create type
 ps <- ps %>% 
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

ps <- as.data.frame(ps)

```

#Private School -skip 

```{r}
# #Subset private schools
#  private <- school10_5 %>% 
#    select("date", "private_school","active_cases_7", "cumulative_faculty_staff_cases_8",                   "cumulative_student_cases_9", "cumulative_total_10")
# #create type
# private <- private %>% 
#    mutate(type="private_schools")
# 
# #rename columns
# colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 
# 
# private <- private %>% 
#   distinct()
# private <- as.data.frame(private)
# private
```

#College

```{r}
#Subset private schools
 college <- school10_5 %>% 
   select("date", "college_university","uni_active_cases", "uni_cumulative_faculty_staff_cases",                   "uni_cumulative_student_cases", "uni_cumulative_total")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school10_5)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school10_5 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","cumulative_active_cases_less_than_5", "cumulative_faculty_staff_cases_less_than_5", "cumulative_student_cases_less_than_5", "cumulative_total_less_than_5")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
"totals_for_all_schools"                            
[22] "cumulative_active_cases_22"                        
[23] "cumulative_faculty_staff_cases_23"                 
[24] "cumulative_student_cases_24"                       
[25] "cumulative_total_25"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school10_5 %>% 
   select("date", "totals_for_all_schools","all_schools_cumulative_active_cases", "all_schools_cumulative_faculty_staff_cases", "all_schools_cumulative_student_cases", "all_schools_cumulative_total")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
#schools_master5 <- smartbind(ps,private,college,less5,totals_all_schools)
schools_master5 <- smartbind(ps,college,less5,totals_all_schools)
head(schools_master5)
schools_master <- smartbind(schools_master5, schools_master)
```

#school10_01

```{r}
school10_1 <- janitor::clean_names(school10_1)
names(school10_1)
```

#Subset tables
#Public School

```{r}
#Rename
school10_1_bak <- school10_1
#Subset public schools
 ps <- school10_1 %>% 
   select("date", "public_school_district","ps_active_cases", "ps_cumulative_faculty_staff_cases",                   "ps_cumulative_student_cases", "ps_cumulative_total")
#create type
 ps <- ps %>% 
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

ps <- as.data.frame(ps)

```

#Private School - skip

```{r}
# #Subset private schools
#  private <- school10_1 %>% 
#    select("date", "private_school","active_cases_7", "cumulative_faculty_staff_cases_8",                   "cumulative_student_cases_9", "cumulative_total_10")
# #create type
# private <- private %>% 
#    mutate(type="private_schools")
# 
# #rename columns
# colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 
# 
# private <- private %>% 
#   distinct()
# private <- as.data.frame(private)
# private
```

#College

```{r}
#Subset private schools
 college <- school10_1 %>% 
   select("date", "college_university","uni_active_cases", "uni_cumulative_faculty_staff_cases",                   "uni_cumulative_student_cases", "uni_cumulative_total")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school10_1)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school10_1 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","cumulative_active_cases_less_than_5", "cumulative_faculty_staff_cases_less_than_5", "cumulative_student_cases_less_than_5", "cumulative_total_less_than_5")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
"totals_for_all_schools"                            
[22] "cumulative_active_cases_22"                        
[23] "cumulative_faculty_staff_cases_23"                 
[24] "cumulative_student_cases_24"                       
[25] "cumulative_total_25"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school10_1 %>% 
   select("date", "totals_for_all_schools","all_schools_cumulative_active_cases", "all_schools_cumulative_faculty_staff_cases", "all_schools_cumulative_student_cases", "all_schools_cumulative_total")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
#schools_master1 <- smartbind(ps,private,college,less5,totals_all_schools)
schools_master1 <- smartbind(ps,college,less5,totals_all_schools)
head(schools_master1)
schools_master <- smartbind(schools_master1, schools_master)
write.csv(schools_master, "schools_master.csv")
```
