---
title: "School Data"
author: "Rob Wells"
date: "10/19/2020"
output: pdf_document
---

# Compilation of School COVID-19 Reports for Arkansascovid.com Calculations

- **This notebook retrieves data from a Google sheet and compiles a time series of ADH school infection reports**

-**It creates a masterfile: schools_master.csv**

# Part 1: Import Data, Clean It

```{r include=FALSE}
#install.packages("slider")
#install.packages("zoo")
#install.packages("gtools")
# vignette("basic-usage", package="googlesheets")
#install.packages("googledrive")
#install.packages("googlesheets4")
library(tidyverse)
library(janitor)
library(lubridate)
library(googlesheets4)
library(googledrive)
library(tidyr)
library(jsonlite)
library(gtools)
library(zoo)  
library(reshape2)
library(slider)

```



#For Googlesheet access
```{r include=FALSE}
#url1 <- ("https://docs.google.com/spreadsheets/d/1UmaSSoMbou2MpvGmihP5cZPNqBnaxqIH60cs10sShB8/edit?usp=sharing")

#school10_15 <- read_sheet(url1, sheet = "Oct15",  col_types = NULL, na = "",trim_ws = TRUE)
#school10_15$date <- "2020-10-15"
school10_12 <- read_sheet(url1, sheet = "Oct12",  col_types = NULL, na = "",trim_ws = TRUE)
school10_8 <- read_sheet(url1, sheet = "Oct8",  col_types = NULL, na = "",trim_ws = TRUE)
school10_5 <- read_sheet(url1, sheet = "Oct5",  col_types = NULL, na = "",trim_ws = TRUE)
school10_1 <- read_sheet(url1, sheet = "Oct1",  col_types = NULL, na = "",trim_ws = TRUE)
school9_28 <- read_sheet(url1, sheet = "Sept28",  col_types = NULL, na = "",trim_ws = TRUE)
school10_12$date <- "2020-10-12"
school10_8$date <- "2020-10-08"
school10_5$date <- "2020-10-05"
school10_1$date <- "2020-10-01"
```


```{r}
#glimpse(school10_15)
```

```{r}
school10_15 <- janitor::clean_names(school10_15)
names(school10_15)
```

#Subset tables
#Public School

```{r}
#Rename
school10_15_bak <- school10_15
#Subset public schools
 ps <- school10_15 %>% 
   select("date", "public_school_district","active_cases_2", "cumulative_faculty_staff_cases_3",                   "cumulative_student_cases_4", "cumulative_total_5")
#create type
 ps <- ps %>% 
   mutate(type="public_schools")

#rename columns
colnames(ps)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

ps <- as.data.frame(ps)

```

#Private School

```{r}
#Subset private schools
 private <- school10_15 %>% 
   select("date", "private_school","active_cases_7", "cumulative_faculty_staff_cases_8",                   "cumulative_student_cases_9", "cumulative_total_10")
#create type
private <- private %>% 
   mutate(type="private_schools")

#rename columns
colnames(private)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

private <- private %>% 
  distinct()
private <- as.data.frame(private)
private
```

#College

```{r}
#Subset private schools
 college <- school10_15 %>% 
   select("date", "college_university","active_cases_12", "cumulative_faculty_staff_cases_13",                   "cumulative_student_cases_14", "cumulative_total_15")
#create type
college <- college %>% 
   mutate(type="college")

#rename columns
colnames(college)[2:6] <-c("name", "active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

college <- college %>% 
  distinct()
college <- as.data.frame(college)
college
```

```{r}
names(school10_15)
```

#schools_with_less_than_5_active_cases

```{r}
#Subset schools_with_less_than_5_active_cases
less5 <- school10_15 %>% 
   select("date", "totals_among_schools_with_less_than_5_active_cases","cumulative_active_cases_17", "cumulative_faculty_staff_cases_18", "cumulative_student_cases_19", "cumulative_total_20")
#create type
less5 <- less5 %>% 
   mutate(type="less5")

#rename columns
colnames(less5)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

less5 <- less5 %>% 
  distinct()
less5 <- as.data.frame(less5)
less5
```

#totals_for_all_schools
"totals_for_all_schools"                            
[22] "cumulative_active_cases_22"                        
[23] "cumulative_faculty_staff_cases_23"                 
[24] "cumulative_student_cases_24"                       
[25] "cumulative_total_25"          
```{r}
#Subset totals_for_all_schools
totals_all_schools <- school10_15 %>% 
   select("date", "totals_for_all_schools","cumulative_active_cases_22", "cumulative_faculty_staff_cases_23", "cumulative_student_cases_24", "cumulative_total_25")
#create type
totals_all_schools<- totals_all_schools %>% 
   mutate(type="totals_all_schools")

#rename columns
colnames(totals_all_schools)[2:6] <-c("name", "cumulative_active_cases", "cumulative_faculty_staff_cases", "cumulative_student_cases", "cumulative_total") 

totals_all_schools <- totals_all_schools %>% 
  distinct()
totals_all_schools <- as.data.frame(totals_all_schools)
totals_all_schools
```

COMBINE USING smartbind
```{r}
schools_master <- smartbind(ps,private,college,less5,totals_all_schools)
head(schools_master)
write.csv(schools_master, "schools_master.csv")
```

#repeat lines 61-174 for each:
school10_12
school10_8
school10_5
school10_1