---
title: "Vaccine Race Data"
output: html_document
---

#ARKANSAS POPULATION BY RACE AND AGE
#SOURCE DATA: https://www.cdc.gov/nchs/nvss/bridged_race/data_documentation.htm#vintage2019
#We downloaded "pcen_v2019_y1019.sas7bdat.zip zip icon[SAS – 58 MB] (Released 7/9/2020)"
#Then carved out the Arkansas portion
#Then matched the county FIPS data for county names
#Then filtered for population age 12+ in all counties
#Then summarized the file to provide total race by age 12+

```{r include=FALSE}
#install.packages("slider")
#install.packages("zoo")
#install.packages("gtools")
#install.packages("kableExtra")
#install.packages("formattable")
library(tidyverse)
library(janitor)
library(lubridate)
library(jsonlite)
library(gtools)
library(zoo)  
library(reshape2)
library(slider)
library(formattable)

```



#REPRODUCIBLE EXAMPLE FROM HERE -- SEE LINES 101-156 ON HOW ARK_RACE WAS CREATED
```{r}

ark_race <- rio::import("https://raw.githubusercontent.com/profrobwells/CovidFall2020/master/ark_race.csv")
ark_race[2:10] <- lapply(ark_race[2:10], as.numeric)

```


#sum race all ages
```{r}
ark_race2 <- ark_race %>% 
  select(County, co_fips, age, hisp, racesex, race4, pop2019, st_fips) %>% 
  group_by(County, race4) %>% 
  summarize(total_pop2019=sum(pop2019))


```


```{r}
#Population totals by race
#Thanks Sean, Kara and Nick:
ark_race2 <- ark_race2 %>%
  mutate(Race = as.character(race4)) %>%   
     mutate(Race = case_when(
    str_detect(Race, "1") ~ "White",
     str_detect(Race, "2") ~ "Black",
     str_detect(Race, "3") ~ "Nat_Am",
     str_detect(Race, "4") ~ "Asian",
    TRUE ~ Race
  ))
```

#sum hispanic all gaes
```{r}
ark_race_hisp <- ark_race %>% 
  select(County, co_fips, age, hisp, racesex, race4, pop2019, st_fips) %>% 
  group_by(County, hisp) %>% 
  summarize(total_pop2019=sum(pop2019)) %>% 
  filter(hisp=="2") 

ark_race_hisp$Race <- str_replace_all(ark_race_hisp$hisp, pattern=fixed('2'), replacement=fixed('Hispanic') )

ark_race_hisp <- ark_race_hisp %>% 
  select(County, total_pop2019, Race)

```

```{r}

x <- ark_race2 %>% 
  select(County, total_pop2019, Race)

Arkansas_Race_Population <- rbind(ark_race_hisp,x)

write.csv(Arkansas_Race_Population, "Arkansas_Race_Population.csv")

```


#Your TURN, Robert Stewart!!
#fact check using the following steps but eliminate the age filter at line 40-41. 
#see if the population totals align with the census 2019 population totals. if they do, then we are in business

#CREATING ARK_RACE

#Import nationwide race and county data
```{r}
sas <- rio::import("/Volumes/1TB SDD MacTower/Wells Data/Dropbox/Classes/Data Fall 2020/ArkansasCovid/Vaccine Race Data/pcen_v2019_y1019.sas7bdat")

sas <- janitor::clean_names(sas)
names(sas)

```

#Subset an Arkansas dataframe from above
```{r}
sas_ark <- sas %>% 
  filter(st_fips =="5")

ark_12_older <- sas_ark %>% 
  filter(age >11)

#select just 2019 data
ark_12_older <- ark_12_older %>% 
  select(age, hisp, racesex, vintage, race4, pop2019, st_fips, co_fips)

```

#join county data
```{r}

counties2 <- rio::import("https://raw.githubusercontent.com/kjhealy/fips-codes/master/state_and_county_fips_master.csv")


ar_counties <- counties %>% 
  filter(state== "AR")

ar_counties$fips1 <- ar_counties$fips
#split data into two columns
ar_counties <- separate(data = ar_counties, col = fips1, into = c("junk", "fips2"), sep = "5", extra = "merge", fill = "right")

#numeric value
ar_counties$fips2 <- as.numeric(ar_counties$fips2)

#merge with ar_12_older

ark_race <- ark_12_older %>% 
  inner_join(ar_counties, by=c("co_fips"="fips2"))

ark_race <- ark_race %>% 
  rename(County = name)

ark_race$County <- gsub(" County", "", ark_race$County)

write.csv(ark_race, "ark_race.csv")

names(ark_race)
```


