---
title: "Arkansas Covid Lesson #2"
author: "Rob Wells"
date: "8/21/2020"
output: html_document
---

# Analyzing Arkansas Covid Data
### Data analysis by Rob Wells


- **The records come from this website**  

https://arkansascovid.com/

Using this sheet:
https://docs.google.com/spreadsheets/d/17M92KbKw1nIOD_co11hN_B0o4AmAW2zs1Nb3QGeDl-I/edit?usp=sharing

--------------------------------------------------------------------
# Part 1: Fix the data structure

```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages("googledrive")
#install.packages("googlesheets4")
library(googlesheets4)
library("googledrive")
library(tidyr)
# vignette("basic-usage", package="googlesheets")
```




```{r}
url <- ("https://docs.google.com/spreadsheets/d/17M92KbKw1nIOD_co11hN_B0o4AmAW2zs1Nb3QGeDl-I/edit#gid=1357388768")

covid <- read_sheet(url, sheet="BeAll")
```

- **Plan B: You may need to copy this Google sheet to your own Google account and then load that URL on to Line 38 above**
https://docs.google.com/spreadsheets/d/17M92KbKw1nIOD_co11hN_B0o4AmAW2zs1Nb3QGeDl-I/edit#gid=1357388768


- **Clean Names**
```{r}

covid <- janitor::clean_names(covid)
```

**Examine the Data Fields**
```{r}
glimpse(covid)
```

- **Extract Population**
```{r}
population <- covid %>% 
  select(county, date, population)
```

- **One Time - Filter Arkansas Counties and One Entry**
```{r}
population_AR <- population %>% 
  filter(county == "Arkansas (All counties)") %>% 
  filter(date > "2020-08-18")
write_csv(population_AR, "population_AR.csv")
```


- **One Time - Filter To One Date For Just Counties and Population**
```{r}
population1 <- population %>% 
  filter(date > "2020-08-18") %>% 
  filter(!(county == "Arkansas (All counties)")) 

population1 <- population1 %>% 
  select(county, population)
  
  write_csv(population1, "population.csv")
```

- **One Time: Extract The Equivalent of the API Feed**
County	Date	Positive	Negative	Recovered	Total Deaths	Number Tested	Current Infections

```{r}
api <- covid %>% 
  select(county, date, positive, negative, recovered, total_deaths, number_tested, current_infections) %>% 
  filter(!(county == "Arkansas (All counties)"))
```


- **Each Day:** We join the new data, api, with the population and then perform the calculations

```{r}
daily <- api %>% 
  # filter(date > "2020-08-18") %>% filtered one time for data validation - all ok
  inner_join(population1, by="county")
head(daily)
```

## MATH!   

- **Now We Do The Math**
- **Previous Date Calculations**

Create Two Days' Worth of Data, Sort Alphbetically and Run Calculations
```{r}
daily1 <- daily %>% 
  filter(date > "2020-08-17") #select YESTERDAY to get two days of data
```

```{r}
daily1 <- daily1 %>% 
  arrange(county) 
```

New Cases = Positive Today - Positive Yesterday
```{r}
daily1 <- daily1 %>% 
  mutate(new_cases_today = (lag(positive)-positive)) %>% 
  mutate(recovered_since_yesterday = (lag(recovered)-recovered)) %>% 
  mutate(new_deaths_today = (lag(total_deaths)-total_deaths)) %>% 
  mutate(new_tests_dashboard = (lag(number_tested)-number_tested)) 
```

Fact Check
Problem: Correct Answer is Displayed on the YESTERDAY FIELD
```{r}
x <- daily1 %>% 
  select(county, date, number_tested, new_tests_dashboard)
x
```

WORKAROUND
BUILD A SEPARATE TABLE THAT SORT BY PREVIOUS DATE
RENAME FIELDS, REJOIN WITH MASTER
```{r}
daily2 <- daily1 %>% 
  select(county, date, new_cases_today, recovered_since_yesterday, new_deaths_today, new_tests_dashboard) %>% 
    filter(date < "2020-08-18") #CAPTURES RESULTS ON THE YESTERDAY LINE
```

RENAMES THE DATE TO TODAY
```{r}
daily2$date <- str_replace_all(daily2$date, pattern=fixed('2020-08-18'), replacement=fixed('2020-08-19') )
```

NORMALIZE DATES BEFORE JOINING
```{r}

daily$date <- as.Date(daily$date) 

daily2$date <- as.Date(daily2$date) 
```

```{r}
glimpse(daily)
```

```{r}
glimpse(daily2)
```
REJOIN WITH DAILY
```{r}
daily_w_cals <- daily %>% 
  inner_join(daily2, by=c("county" = "county", "date" = "date"))
```


- **Simplify: Get a subset table**
```{r}
cases <- covid %>% 
  select(county, date, current_infections, new_cases_today)
head(cases)
```

- **Filter: Just the state totals**

```{r}
state <- cases %>% 
  select(county, date, current_infections, new_cases_today) %>% 
  filter(county=="Arkansas (All counties)") 
head(state)
```

- **Graph state totals**

```{r}
ggplot(state, aes(x = date, y = new_cases_today, fill=new_cases_today)) + 
  geom_bar(stat = "identity", show.legend = FALSE) +
  #geom_text(aes(label = new_cases_today), hjust = -.1, size = 2.2) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Arkansas: New Covid-19 Cases Per Day", 
       subtitle = "Source: Arkansascovid.com",
       caption = "Graphic by Wells",
       y="New cases per day",
       x="Date")

#ggsave("/Users/rswells/Dropbox/Classes/Data Fall 2020/CovidFall2020/Xx.png",device = "png",width=9,height=6, dpi=800)
```

- **Your Turn**

      Build a table with deaths by state 
      Build a table with new cases and deaths by state
      Build a table with news cases and deaths by Washington County


---

- **Lesson #2**

--Based on this tutorial, perform this exercise:  

    1. Percentage change from March-August.
    2. Produce a table with 5 counties with most growth in cases. 
    3. Graph the top 5 and bottom 5
    4. Filter just Benton County new cases and deaths for July
    
    
#Edit Below


#Get census data
#Check the website: 
https://factfinder.census.gov/bkmk/table/1.0/en/PEP/2017/PEPANNRES/0400000US05.05000


#Load table - ArkCensus.csv - 
download.file("https://bit.ly/2FxLJHj", "ArkCensus.csv")

#It downloaded it to your current working directory. Use this command to find where it is
getwd() 

#Import using Rio. Make sure ArkCensus is in your working directory!
ArkCensus <- rio::import("ArkCensus.csv")

#Look at the table
View(ArkCensus)

# another method Rename a specific column
colnames(ArkCensus)[5] <- "BaseEstimate"
View(ArkCensus)


#Create a New Column and a Formula: Percents of Washington County as Whole of the state of Arkansas

ArkCensus$Pct2017 <- ((ArkCensus$x2017-ArkCensus$x2016)/(ArkCensus$x2016))

#To quickly format into percents, load
install.packages("formattable")
library(formattable)

ArkCensus$Pct2017 <- percent(ArkCensus$Pct2017)

View(ArkCensus)


# Sort to see biggest-smallest descending in population from 2016-2017 

ArkCensus <- ArkCensus[order(-ArkCensus$Pct2017),]
View(ArkCensus)

# What is the average population change?

mean(ArkCensus$Pct2017)
median(ArkCensus$Pct2017)

#Find all counties with population growth above 1%

Above1pct <- ArkCensus%>%select(county, Pct2017)%>%filter(Pct2017 > .01)

#Build a table with places with upper quantile of crime
quantile(ArkCensus$Pct2017)

# the Upper Quantile is 0.74% to 2.94%
TopGrowth <- ArkCensus%>%select(county, Pct2017)%>%filter(Pct2017 > 0.0074)

#Find all places with below average crime

MajorLosers <- ArkCensus%>%select(county, Pct2017)%>%filter(Pct2017 < -0.0093)

#Filter
Filter
Washington1 <- subset(ArkCensus, County=="Washington County, Arkansas")  

Using Wildcards
WashingtonCo <- subset(ArkCensus, grepl("^Washington", County))


#Write.csv
write.csv(Washington1, "Washington1.csv")


#-------------------------------------------------------------------#
#Build a chart - Total loan disbursement
#-------------------------------------------------------------------#

library(ggplot2)

TopGrowthChart <- ggplot(TopGrowth, aes(x = county, y=Pct2017)) +
geom_bar(stat = "identity") +
coord_flip() +     #this makes it a horizontal bar chart instead of vertical
labs(title = "Your title here", 
subtitle = "Subtitle and source ",
caption = "Graphic by Rob Wells",
x="County",
y="Population growth 2016-2017")
plot(TopGrowthChart)


#-------------------------------------------------------------------#
#      What You Have Learned So Far
#-------------------------------------------------------------------#  

# How to navigate in R studio
# How to install libraries and packages 
# How to import a .csv file into R: read.csv
# How to obtain summary statistics (summary)
# How to create a new calculated field

