---
title: "Maps Demos 6th Lesson 8-22-2020"
author: "Rob Wells"
date: "8/22/2020"
output: html_document
---

	demo(webmap)
	---- ~~~~~~

```{r} 
library(sp)
```

```{r} 
library(ggmap)
```
Loading required package: ggplot2
Google's Terms of Service: https://cloud.google.com/maps-platform/terms/.
Please cite ggmap if you use it! See citation("ggmap") for details.

```{r} 
demo(meuse, ask=FALSE, echo = FALSE)
```

```{r} 
merc = CRS("+init=epsg:3857")
```

```{r} 
WGS84 = CRS("+init=epsg:4326")
```

```{r} 
meuse = spTransform(meuse, WGS84)
```

```{r} 
bgMap = get_map(as.vector(bbox(meuse)), source = "google", zoom = 13) # useless without zoom level
```
Source : http://tile.stamen.com/terrain/13/4226/2742.png
Source : http://tile.stamen.com/terrain/13/4227/2742.png
Source : http://tile.stamen.com/terrain/13/4226/2743.png
Source : http://tile.stamen.com/terrain/13/4227/2743.png
Source : http://tile.stamen.com/terrain/13/4226/2744.png
Source : http://tile.stamen.com/terrain/13/4227/2744.png

# plot with ggmap-google bg:
```{r} 
plot(spTransform(meuse, merc), bgMap = bgMap, pch = 16, cex = .5)
```

# spplot with ggmap-google bg:
```{r} 
spplot(spTransform(meuse, merc), c("zinc",  "lead"), colorkey = TRUE,
+ 	sp.layout = list(panel.ggmap, bgMap, first = TRUE))
```
# plot with ggmap-osm bg:
```{r} 
bb = t(apply(bbox(meuse), 1, bbexpand, .04))
```

```{r} 
bgMap = get_map(as.vector(bb), source = "osm") # WGS84 for background map
```
Source : http://tile.stamen.com/terrain/15/16904/10971.png
Source : http://tile.stamen.com/terrain/15/16905/10971.png
Source : http://tile.stamen.com/terrain/15/16906/10971.png
Source : http://tile.stamen.com/terrain/15/16907/10971.png
Source : http://tile.stamen.com/terrain/15/16908/10971.png
Source : http://tile.stamen.com/terrain/15/16904/10972.png
Source : http://tile.stamen.com/terrain/15/16905/10972.png
Source : http://tile.stamen.com/terrain/15/16906/10972.png
Source : http://tile.stamen.com/terrain/15/16907/10972.png
Source : http://tile.stamen.com/terrain/15/16908/10972.png
Source : http://tile.stamen.com/terrain/15/16904/10973.png
Source : http://tile.stamen.com/terrain/15/16905/10973.png
Source : http://tile.stamen.com/terrain/15/16906/10973.png
Source : http://tile.stamen.com/terrain/15/16907/10973.png
Source : http://tile.stamen.com/terrain/15/16908/10973.png
Source : http://tile.stamen.com/terrain/15/16904/10974.png
Source : http://tile.stamen.com/terrain/15/16905/10974.png
Source : http://tile.stamen.com/terrain/15/16906/10974.png
Source : http://tile.stamen.com/terrain/15/16907/10974.png
Source : http://tile.stamen.com/terrain/15/16908/10974.png
Source : http://tile.stamen.com/terrain/15/16904/10975.png
Source : http://tile.stamen.com/terrain/15/16905/10975.png
Source : http://tile.stamen.com/terrain/15/16906/10975.png
Source : http://tile.stamen.com/terrain/15/16907/10975.png
Source : http://tile.stamen.com/terrain/15/16908/10975.png
Source : http://tile.stamen.com/terrain/15/16904/10976.png
Source : http://tile.stamen.com/terrain/15/16905/10976.png
Source : http://tile.stamen.com/terrain/15/16906/10976.png
Source : http://tile.stamen.com/terrain/15/16907/10976.png
Source : http://tile.stamen.com/terrain/15/16908/10976.png

```{r} 
plot(spTransform(meuse, merc), bgMap = bgMap, pch = 16, cex = .5)
```
# RgoogleMaps:
```{r} 
center = apply(coordinates(meuse), 2, mean)[2:1]
```

```{r} 
library(RgoogleMaps)
```

```{r} 
g = GetMap(center=center, zoom=13) # google
```

```{r} 
par(mar = rep(0,4)) # fill full device
```

```{r} 
plot(spTransform(meuse, merc), bgMap = g, pch = 16, cex = .5)
```

```{r} 
spplot(spTransform(meuse, merc), c("zinc",  "lead"), colorkey = TRUE,
+ 	sp.layout = list(panel.RgoogleMaps, g, first = TRUE),
+ 	scales = list(draw = TRUE))
```

# Norway boundary example:
```{r} 
library(cshapes)
```