---
title: "Math Intro 4th Lesson 8-21-2020"
author: "Rob Wells"
date: "8/21/2020"
output: html_document
---

# Math 

```{r}
library(tidyverse)
StudentLoans <- rio::import('https://raw.githubusercontent.com/profrobwells/CovidFall2020/master/Data/AR2016_SMALL.csv')
View(StudentLoans)
```

Cut it down to size

```{r}
StLoans2 <- StudentLoans %>% 
  select(INSTNM, CITY, UGDS_WHITE, UGDS_BLACK)
head(StLoans2)
```


- **Summary Statistics**
    
    A quick way to view the range of your data  

```{r}
summary(StLoans2$UGDS_WHITE)
```

- **Your Turn**

    Run individual calculations using the following operators
    mean(x) Calculate the mean, or average, for variable x.
    median(x) Calculate the median.
    max(x) Find the maximum value.
    min(x) Find the minimum value.
    sum(x) Add all the values together.
    n() Count the number of records. 
    n_distinct(x) Count the number of unique values in variable x.
  
```{r}

  your code here(
    
```
  
Good R Tutorial with Basic Statistics
https://www.princeton.edu/~otorres/sessions/s2r.pdf


- **Create a new average column **   

```{r}
StLoans2$Combo <- (StLoans2$UGDS_WHITE + StLoans2$UGDS_BLACK) / 2 
head(StLoans2)
```
  
- **Using a function for an equation**

```{r}
  percent_change <- function(first_number, second_number) {
  pc <- (second_number-first_number)/first_number*100
  return(pc)
}
```

```{r}
percent_change(100,150)
```

- **What Happened**

      * percent_change is the name of the function, and assigned to it is the function function()
      * Two variables are necessary to be passed to this function, first_number and second_number
      * A new object pc is created using some math calculating percent change from the two variables passed to it
      * the function return() assigns the result of the math to percent_change from the first line
      Build enough functions and you can save them as your own package.


- **Doing math on columns with missing values**

```{r}
sum(StudentLoans$UGDS_WHITE, na.rm=TRUE)
```
```{r}
mean(StudentLoans$TUITIONFEE_IN, na.rm=TRUE)
```



- **Set up column for math calculations**   
      Example: Total column shows winter snowfall in inches. To add a column    showing totals in Meters, you can use this format:   

```{r}      
snowdata$Meters <- snowdata$Total * 0.0254
```


**--30--**